---
title: "Lambda Calculations of Phylogenetic Signal"
output: html_notebook
---

# Introduction

This notebook serves to help prepare the tree to be used in this analysis and run and visualise the output of the phylogenetic signal analysis. Hence, this notebook creates Figure 3 (dot plot of ranked lambda values for continous traits; exported filename 'lambda_values_dotplot.pdf') and supplementary Table S3.1 (comparison of lambda values of 15-17 different possible trees to assess effect of different ambiguous taxon positions on analysis; exported at the end of script 3a) in the manuscript.

The statistical analysis was run by the `pgls()` function from the package 'caper' (Orme et al. 2018).

# Analysis

## Required Packages, Functions, and Data


```{r}
# packages

library(ape)
library(ggtree)
library(caper)
library(ggplot2)



# functions

## shortcut fot plotting trees (of the relevant size for this analysis) with the right dimensions to be easily inspected in the output within the this markdown document
plot_dim1 <- function(phy) {
  # plots with the correct size text and alignment for 158-225 taxa trees like will be used
  plot(ggtree(phy) +xlim(0, 35) + geom_tiplab(size=5)) 
}

## the following shortcut works better regarding parameters of plotting trees in some cases where plot_dim1 shrinks the tree severely to the left
plot_dim2 <- function(phy) {
  # plots with the correct size text and alignment for 158-225 taxa trees like will be used
  plot(ggtree(phy) +xlim(0, 1.5) + geom_tiplab(size=5)) 
}

# The function below is required for extracting the boundaries of the confidence interval for lambda (so it can be plotted with its error bars) and separates the boundaries of the confidence interval of parameters (lambda) into two numeric objects instead of a single object

divide_ci <- function(ci.val) { 
  ## requires input of class 'list' of confidence intervals separated by ' ' (space), since that is how the pgls model object stores that info
  ## classed as array
  bounds.long <- stringr::str_split_fixed(ci.val, ' ', 2) 
  ## creating an empty data frame to then populate with values in the next step
  bounds.split <- data.frame('lower_CI' = rep(NA, 10), 'upper_CI' = rep(NA, 10))
  for (i in 1:10) {
    ## subsetting arrays (rather than data frames /lists allows the following to mean only the first column if no column is specified
    bounds.split[i, c('lower_CI')] <- as.numeric(bounds.long[2*i - 1]) 
    bounds.split[i, c('upper_CI')] <- as.numeric(bounds.long[2*i])
  }
  return(bounds.split)
}

# data
eucs.data <- read.csv("../Input_data/eucs_data.csv")
eucs.log.data <- read.csv("../Input_data/eucs_log_data.csv")
eucs.scaled.data <- read.csv("../Input_data/eucs_scaled_data.csv")

# phylogenetic tree
tree <- read.tree("../Input_data/final_tree.txt")
```

## Data Preparation

```{r fig.width = 15, fig.height = 25}
# First to get the imported phylogeny into correct form for joining with trait data
## remove some erroneous node labels that cause issues later in case they appear
tree$node.label <- NULL
## converts tip names back into space-separated rather than underscore (the default when reading from newick format)
tree$tip.label <- gsub("_", " ", tree$tip.label)
## trees require branch length for analysis, this method generates it from topology-only trees allowing us to calculate genetic closeness beween taxa based on the number of nodes from the root they share instead of using the length of branches, which has been removed as part of the tree-building process anyway
tree <- compute.brlen(tree, method = "Grafen", power = 1)
# just to check that the tree looks normal and like what we expect it to
plot_dim2(tree)


# Preparing the matching objects that allow the package to link data frames to tree tips

## for the untransformed data in case we want to use it
eucs.obj <- comparative.data(phy = tree, data = eucs.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)
eucs.obj$dropped$unmatched.rows # the only unmatched taxa are: E. bunyip and E. carolaniae, E. dendromorpha and E. robusta, and Tristaniopsis laurina
eucs.obj$dropped$tips #none as no tips on the tree are redundant

## for the log-transformed data in case we want to use it
eucs.log.obj <- eucs.log.obj <- comparative.data(phy = tree, data = eucs.log.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)
eucs.log.obj$dropped$unmatched.rows 

## for the scaled data, which we will most likely use
eucs.scaled.obj <- comparative.data(phy = tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)
eucs.scaled.obj$dropped$unmatched.rows
```


## Phylogenetic Signal using Pagel's lambda

Now to first compute lambda for each trait before testing correlations.

```{r}
lambda.mh <- pgls(max_height_m ~ 1, data = eucs.scaled.obj, lambda = 'ML', bounds = list(lambda = c(0.0001, 1), kappa = c(1e-6, 3), delta = c(1e-6, 3)))
summary(lambda.mh)
```

The intercept for the scaled data is not meant to be significantly different from zero bc scaled data by definition has a mean of 0 anyway. However, the phylogenetic mean is not the same as the equally weighted mean across the data, and tends to represent what the trait value is estimated to be at the root node of the whole tree.

```{r}
lambda.rh <- pgls(relative_height_by_girth ~ 1, data = eucs.scaled.obj, lambda = 'ML')
summary(lambda.rh)
```


```{r}
lambda.std <- pgls(stem_density_g_per_ml ~ 1, data = eucs.scaled.obj, lambda = 'ML')
summary(lambda.std)
```


```{r}
lambda.rbt <- pgls(relative_bt_by_girth ~ 1, data = eucs.scaled.obj, lambda = 'ML')
summary(lambda.rbt)
```


```{r}
lambda.sla <- pgls(sla_mm2_per_mg ~ 1, data = eucs.scaled.obj, lambda = 'ML', bounds = list(lambda = c(0.0001, 1), kappa = c(1e-6, 3), delta = c(1e-6, 3)))
summary(lambda.sla)
```


```{r}
lambda.lfa <- pgls(leaf_area_cm2 ~ 1, data = eucs.scaled.obj, lambda = 'ML')
summary(lambda.lfa)
```


```{r}
lambda.lms <- pgls(leaf_mass_g ~ 1, data = eucs.scaled.obj, lambda = 'ML', bounds = list(lambda = c(0.001, 1), kappa = c(1e-6, 3), delta = c(1e-6, 3)))
summary(lambda.lms)
```


```{r}
lambda.fms <- pgls(fruit_mass_mg ~ 1, data = eucs.scaled.obj, lambda = 'ML')
summary(lambda.fms)
```


```{r}
lambda.fwl <- pgls(fruit_wall_width_mm ~ 1, data = eucs.scaled.obj, lambda = 'ML')
summary(lambda.fwl)
```


```{r}
lambda.sms <- pgls(seed_mass_mg ~ 1, data = eucs.scaled.obj, lambda = 'ML', bounds = list(lambda = c(0.0001, 1), kappa = c(1e-6, 3), delta = c(1e-6, 3)))
summary(lambda.sms)
```



# Dot-Plot of Lambda values

```{r}
# First to create a data frame of lambda values of each model corresponding to each trait
lambda.full <- data.frame('Traits' = c("Max. height", "Relative Height", "Stem Density", "Relative Bark Thickness", "SLA", "Leaf Area", "Leaf Mass", "Fruit Mass", "Fruit Wall Width", "Seed Mass"), 'Lambda' =  c(lambda.mh$param.CI$lambda$opt, lambda.rh$param.CI$lambda$opt, lambda.std$param.CI$lambda$opt, lambda.rbt$param.CI$lambda$opt, lambda.sla$param.CI$lambda$opt, lambda.lfa$param.CI$lambda$opt, lambda.lms$param.CI$lambda$opt, lambda.fms$param.CI$lambda$opt, lambda.fwl$param.CI$lambda$opt, lambda.sms$param.CI$lambda$opt))

# Extract the bound of the confidence interval around the lambda value each time
ci.val <- c(lambda.mh$param.CI$lambda$ci.val, lambda.rh$param.CI$lambda$ci.val, lambda.std$param.CI$lambda$ci.val, lambda.rbt$param.CI$lambda$ci.val, lambda.sla$param.CI$lambda$ci.val, lambda.lfa$param.CI$lambda$ci.val, lambda.lms$param.CI$lambda$ci.val, lambda.fms$param.CI$lambda$ci.val, lambda.fwl$param.CI$lambda$ci.val, lambda.sms$param.CI$lambda$ci.val)

# divides the bounds of the 95% CI into two separate objects, one for the lower bounds and another for the upper bound using custom function
bounds <- divide_ci(ci.val)

# now to create full data frame
lambda.full <- cbind(lambda.full, bounds)
## since the likely reason for the NA value next to relative height is that it's non-significance would take the lower bound below 0, which is technically below what values lambda can take (or is tested for by this function), hence for our purposes of calculating range and plotting error bars it is fine to replace it with 0 for now
lambda.full[lambda.full$Traits == "Relative Height", c("lower_CI")] <- 0 
## having an extra column to compare ranges/widths of confidence intervals
lambda.full$range_CI <- lambda.full$upper_CI - lambda.full$lower_CI
# export this table so it can be used to add things to suppmat tables comparing the lambda values and ranges of values for potential trees to the value and range found for this final tree
write.csv(lambda.full, file = "../Output_figures/final_tree_lambdas.csv")

## for dot plot, need categorical traits to be in order from highest to lower lambda value
dot.plot.data <- lambda.full[order(lambda.full$Lambda),]

# the following plots the dotplot with parameters such that it is most easily visible when exported as a pdf (i.e. how it will be in the final manuscript). However, these parameters cause much overlap when attempting to view the plot within this Rmarkdown, so exporting before viewing is recommended
lambda.dot.plot <- ggplot(dot.plot.data, aes(x=Traits, y=Lambda), dotsize = 3) + theme_classic() + geom_dotplot(binaxis='y', stackdir='center') + scale_x_discrete(limits=dot.plot.data$Traits) + 
scale_y_continuous(limits = c(0, 1)) + geom_errorbar(aes(ymin=lower_CI, ymax=upper_CI, width=.1)) + theme(axis.text.y = element_text(size = 30), axis.title.x = element_text(size = 30), axis.text.x = element_text(size = 30)) +labs(y= "Pagels Lambda", x="")+ coord_flip()

# to save the dot plot
ggsave(lambda.dot.plot, file = "../Output_figures/lambda_values_dotplot.pdf", width = 15, height = 10)
```

All values but that calculated for Relative height were significantly different from zero, hence there is detectable phylogenetic signal in all the traits but that one. Leaf mass and Relative Bark Thickness had values significantly different from zero, but it was very low. The rest had moderate to high values of lambda.

Relative height may not be a good trait at all and may exhibit greater variation between individuals or populations of the same taxon than between taxa themselves, i.e. potentially also not a stable trait within a taxon.
