---
title: "Lambda Calculations of Phylogenetic Signal"
output: html_notebook
---

# Introduction

This notebook serves to prepare each of the 15-17 possible candidate trees created in notebook 2 (section 'Testing Placement Effects') and run and visualise the output of the phylogenetic signal analysis (i.e. calculation of Pagel's Lambda) so that differences in those outputs can be compared. Upon selection of a final tree to be used in the main and subsequent analyses, this notebook combines the output of Lambda value calculation of this final tree with that of all the possible candidate trees (created by notebook/script 3) in a data frame subsequently used for supplementary Table S3.1 of the manuscript (comparison of lambda values of 15-17 different possible trees to assess effect of different ambiguous taxon positions on analysis; exported filename 'suppmat_pot_tree_lambdas2_vertical_rounded.csv').

The statistical analysis was run by the `pgls()` function from the package 'caper' (Orme et al. 2018).

## How to use this script

A large part of this script acts like one large single function to be rerun for each potential phylogeny. Each time the sections down to the end of 'Dot-Plot of Lambda values' are run (until line 203), the following lines must be changed: (1) choose a different number to put in line 77 to select a different tree from that list of 15 potential trees; (2) change the name of the output figure (line 202) to match the number of the tree in that list so a new plot will be created and not overwrite the previous one; and (3) change the number at the end of the final object named `lambda.full` (but only the instance where it appears in line 195) to the corresponding object name with that number (all are listed in line 217). Rerunning this process for each tree in the list generates and stores the result for each tree run as an object in the environment. This can then be subsequently combined into a data frame (lines 211-237) before being later added to outputs from other scripts (run in the order specified in the README document for this project) to create a combined data frame for exporting (lines 655-end). Lines 240-649 define a section in the middle of this script where the effects of differences between trees on the SLA-LA and MH-SD trait correlation strength are tested and can be observed within this notebook, though no output is exported from this section.

# Analysis

## Required Packages, Functions, and Data


```{r}
# packages

library(ape)
library(ggtree)
library(caper)
library(ggplot2)
library(dplyr)

# functions

## shortcut fot plotting trees (of the relevant size for this analysis) with the right dimensions to be easily inspected in the output within this markdown document
plot_dim1 <- function(phy) {
  # plots with the correct size text and alignment for 158-225 taxa trees like will be used
  plot(ggtree(phy) +xlim(0, 35) + geom_tiplab(size=5)) 
}


# The function below is required for extracting the boundaries of the confidence interval for lambda (so it can be plotted with its error bars) and separates the boundaries of the confidence interval of parameters (lambda) into two numeric objects instead of a single object

divide_ci <- function(ci.val) { 
  ## requires input of class 'list' of confidence intervals separated by ' ' (space), since that is how the pgls model object stores that info
  ## classed as array
  bounds.long <- stringr::str_split_fixed(ci.val, ' ', 2) 
  ## creating an empty data frame to then populate with values in the next step
  bounds.split <- data.frame('lower_CI' = rep(NA, 10), 'upper_CI' = rep(NA, 10))
  for (i in 1:10) {
    ## subsetting arrays (rather than data frames /lists allows the following to mean only the first column if no column is specified
    bounds.split[i, c('lower_CI')] <- as.numeric(bounds.long[2*i - 1]) 
    bounds.split[i, c('upper_CI')] <- as.numeric(bounds.long[2*i])
  }
  return(bounds.split)
}

# rounding function to three decimal places
num_round <- function(x){
  x_num <- as.numeric(x)
  out <- round(x_num, 3)
  return(out)
}


# data
eucs.data <- read.csv("../Input_data/eucs_data.csv")
eucs.log.data <- read.csv("../Input_data/eucs_log_data.csv")
eucs.scaled.data <- read.csv("../Input_data/eucs_scaled_data.csv")
```


## Data Preparation

```{r fig.height = 30, fig.width = 15}
# Each subsequent candidate input tree will be named eucs.test.tree, each tree is known to be rooted in advance
## first to read in the list object of all the newick forms of the 15 potential trees
pot.tree.list <- readRDS("../Intermediate_Tree_Building_Products/Potential_trees_nwk_list.RData")
## picking which potential tree we're going to use for this run
tree <- read.tree(text = pot.tree.list[[15]])
## remove some erroneous node labels that cause issues later
tree$node.label <- NULL
## converts tip names back into space-separated rather than underscore (the default when reading from newick format)
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
## trees require branch length for analysis, this method generates it from topology-only trees allowing us to calculate genetic closeness beween taxa based on the number of nodes from the root they share instead of using the length of branches, which has been removed as part of the tree-building process anyway
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
# plot_dim1(tree)


# Preparing the matching objects that allow the package to link data frames to tree tips

## for the untransformed data
eucs.test <- comparative.data(phy = eucs.test.tree, data = eucs.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)
### the only unmatched taxa dropped from the dataset due to having no match on the tree are some subset of chapmaniana, punctata, silvestris, strzleckii, behriana and wimmerensis
eucs.test$dropped$unmatched.rows 
# Should have no tips dropped as only taxa in the trait dataset should be on the tree
eucs.test$dropped$tips 

## for the log-transformed data
eucs.log.obj <- eucs.log.obj <- comparative.data(phy = eucs.test.tree, data = eucs.log.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)
### following should be same list as for untransformed data `eucs.test` above
eucs.log.obj$dropped$unmatched.rows 

## for the scaled data
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)
### following should be same list as for untransformed data `eucs.test` above
eucs.scaled.obj$dropped$unmatched.rows
```


## Phylogenetic Signal using Pagel's lambda

Now to first compute lambda for each trait before testing correlations.

```{r}
lambda.mh <- pgls(max_height_m ~ 1, data = eucs.scaled.obj, lambda = 'ML', bounds = list(lambda = c(0.0001, 1), kappa = c(1e-6, 3), delta = c(1e-6, 3)))
summary(lambda.mh)
```

The intercept for the scaled data is not meant to be significantly different from zero bc scaled data by definition has a mean of 0 anyway. However, the phylogenetic mean is not the same as the equally weighted mean across the data, and tends to represent what the trait value is estimated to be at the root node of the whole tree.

```{r}
lambda.rh <- pgls(relative_height_by_girth ~ 1, data = eucs.scaled.obj, lambda = 'ML')
summary(lambda.rh)
```


```{r}
lambda.std <- pgls(stem_density_g_per_ml ~ 1, data = eucs.scaled.obj, lambda = 'ML')
summary(lambda.std)
```


```{r}
lambda.rbt <- pgls(relative_bt_by_girth ~ 1, data = eucs.scaled.obj, lambda = 'ML')
summary(lambda.rbt)
```


```{r}
lambda.sla <- pgls(sla_mm2_per_mg ~ 1, data = eucs.scaled.obj, lambda = 'ML', bounds = list(lambda = c(0.0001, 1), kappa = c(1e-6, 3), delta = c(1e-6, 3)))
summary(lambda.sla)
```


```{r}
lambda.lfa <- pgls(leaf_area_cm2 ~ 1, data = eucs.scaled.obj, lambda = 'ML')
summary(lambda.lfa)
```


```{r}
lambda.lms <- pgls(leaf_mass_g ~ 1, data = eucs.scaled.obj, lambda = 'ML', bounds = list(lambda = c(0.001, 1), kappa = c(1e-6, 3), delta = c(1e-6, 3)))
summary(lambda.lms)
```


```{r}
lambda.fms <- pgls(fruit_mass_mg ~ 1, data = eucs.scaled.obj, lambda = 'ML')
summary(lambda.fms)
```


```{r}
lambda.fwl <- pgls(fruit_wall_width_mm ~ 1, data = eucs.scaled.obj, lambda = 'ML')
summary(lambda.fwl)
```


```{r}
lambda.sms <- pgls(seed_mass_mg ~ 1, data = eucs.scaled.obj, lambda = 'ML', bounds = list(lambda = c(0.0001, 1), kappa = c(1e-6, 3), delta = c(1e-6, 3)))
summary(lambda.sms)
```



# Dot-Plot of Lambda values

```{r}
# First to creat a dataframe of lambda values of each model corresponding to each trait
lambda.full <- data.frame('Traits' = c("Max. height", "Relative Height", "Stem Density", "Relative Bark Thickness", "SLA", "Leaf Area", "Leaf Mass", "Fruit Mass", "Fruit Wall Width", "Seed Mass"), 'Lambda' =  c(lambda.mh$param.CI$lambda$opt, lambda.rh$param.CI$lambda$opt, lambda.std$param.CI$lambda$opt, lambda.rbt$param.CI$lambda$opt, lambda.sla$param.CI$lambda$opt, lambda.lfa$param.CI$lambda$opt, lambda.lms$param.CI$lambda$opt, lambda.fms$param.CI$lambda$opt, lambda.fwl$param.CI$lambda$opt, lambda.sms$param.CI$lambda$opt))

# extracts the two values for boundaries of the 95% confidence interval of lamba for each model
ci.val <- c(lambda.mh$param.CI$lambda$ci.val, lambda.rh$param.CI$lambda$ci.val, lambda.std$param.CI$lambda$ci.val, lambda.rbt$param.CI$lambda$ci.val, lambda.sla$param.CI$lambda$ci.val, lambda.lfa$param.CI$lambda$ci.val, lambda.lms$param.CI$lambda$ci.val, lambda.fms$param.CI$lambda$ci.val, lambda.fwl$param.CI$lambda$ci.val, lambda.sms$param.CI$lambda$ci.val)

# splits the CI bounds into two seperate objects, one for the lower and another for the upper bound
bounds <- divide_ci(ci.val)

# constructing the full dataframe of estimate and respective bounds
lambda.full <- cbind(lambda.full, bounds)
## since the likely reason for the NA value next to relative height is that it's non-significance would take the lower bound below 0, which is technically below what values lambda can take (or is tested for by this function), hence for our purposes of calculating range and plotting error bars it is fine to replace it with 0 for now
lambda.full[lambda.full$Traits == "Relative Height", c("lower_CI")] <- 0 
## having an extra column to compare ranges
lambda.full$range_CI <- lambda.full$upper_CI - lambda.full$lower_CI
## for dot plot, need categorical traits to be in order from highest to lower lambda value
dot.plot.data <- lambda.full[order(lambda.full$Lambda),]
lambda.full.null <- lambda.full

# now to plot the tree using the above ordered data frame
lambda.dot.plot <- ggplot(dot.plot.data, aes(x=Traits, y=Lambda), dotsize = 3) + theme_classic() + geom_dotplot(binaxis='y', stackdir='center') + scale_x_discrete(limits=dot.plot.data$Traits) + 
scale_y_continuous(limits = c(0, 1)) + geom_errorbar(aes(ymin=lower_CI, ymax=upper_CI, width=.1)) + theme(axis.text.y = element_text(size = 30), axis.title.x = element_text(size = 30), axis.text.x = element_text(size = 30)) +labs(y= "Pagels Lambda", x="")+ coord_flip()

# now to save each iteration to a seperate pdf
ggsave(lambda.dot.plot, file = "../Output_figures/Potential_trees_and_results/lambda_values_dotplot_null.pdf", width = 15, height = 10)
```

All values but that calculated for Relative height were significantly different from zero, hence there is detectable phylogenetic signal in all the traits but that one. Leaf mass had a value significantly different from zero, but it was very low. However, since updating the dataset in 2022, the lowest lambda value was from relative bark thickness. The rest had moderate to high values of lambda.

Relative height may not be a good trait at all and may exhibit greater variation between individuals or populations of the same taxon than between taxa themselves, i.e. potentially also not a stable trait within a taxon. Perhaps analysis of variance might help with figuring this out??



# Lambda Values Comparison for Above Tests

Now to determine what the maximum range of lambda values are for each trait across all 15 potential trees and compare that to the size of the confidence interval for each trait's lambda value (both the CI of the single tree we will be using from later, as well as the one with the widest range among these 15).

```{r}
# data frame of lambda value estimated for each trait for each of the 15 trees
compare.lambda.trees <- data.frame('Traits' = lambda.full.null$Traits, 'Lambda1' = lambda.full1$Lambda, 'Lambda2' = lambda.full2$Lambda, 'Lambda3' = lambda.full3$Lambda, 'Lambda4' = lambda.full4$Lambda, 'Lambda5' = lambda.full5$Lambda, 'Lambda6' = lambda.full6$Lambda, 'Lambda7' = lambda.full7$Lambda, 'Lambda8' = lambda.full8$Lambda, 'Lambda9' = lambda.full9$Lambda, 'Lambda10' = lambda.full10$Lambda, 'Lambda11' = lambda.full11$Lambda, 'Lambda12' = lambda.full12$Lambda, 'Lambda13' = lambda.full13$Lambda, 'Lambda14' = lambda.full14$Lambda, 'Lambda_null' = lambda.full.null$Lambda)

# adding the column stating the range between the largest and smallest lambda values for a given trait
compare.lambda.trees$range <- unlist(lapply(as.data.frame(t(as.matrix(compare.lambda.trees[,!names(compare.lambda.trees) %in% c('Traits')]))), function(lambdas_for_a_trait){
  range <- range(lambdas_for_a_trait)[[2]] - range(lambdas_for_a_trait)[[1]]
  return(range)
}))


# width of CI for each trait for each of the 15 trees
compare.range.CI.trees <- data.frame('Traits' = lambda.full.null$Traits, 'rangeCI_1' = lambda.full1$range_CI, 'rangeCI_2' = lambda.full2$range_CI, 'rangeCI_3' = lambda.full3$range_CI, 'rangeCI_4' = lambda.full4$range_CI, 'rangeCI_5' = lambda.full5$range_CI, 'rangeCI_6' = lambda.full6$range_CI, 'rangeCI_7' = lambda.full7$range_CI, 'rangeCI_8' = lambda.full8$range_CI, 'rangeCI_9' = lambda.full9$range_CI, 'rangeCI_10' = lambda.full10$range_CI, 'rangeCI_11' = lambda.full11$range_CI, 'rangeCI_12' = lambda.full12$range_CI, 'rangeCI_13' = lambda.full13$range_CI, 'rangeCI_14' = lambda.full14$range_CI, 'rangeCI_null' = lambda.full.null$range_CI)

# adding the smallest CI-range for each trait across all 15 trees to the first data frame to show it is always an order of magnitude greater than the variation in lambda point estmates from tree to tree
compare.lambda.trees$min_CI_range <- unlist(lapply(as.data.frame(t(as.matrix(compare.range.CI.trees[,!names(compare.range.CI.trees) %in% c('Traits')]))), min))
```

It can clearly be seen that even for the tree with the smallest and narrowest CI for the lambda for each trait is an order of magnitude greater than the range of all point estimates of lambda across the 15 trees for a given trait. Once we have decided on the final tree to be used, the CI bounds and CI range for that will also be added as a final 2 columns to confirm that all 15 trees' values do fall within our final CI as well.

Hence, the exact positioning of the few taxa let over for which the true position on the phylogeny is unclear is not largely consequential for the calculation of lambda.

Now to check if the same is true for two of the our most sensitive correlations to lambda value.


# Some Sensitive Correlations

Since the lambda calcuated for a single trait in an intercept-only model is not necessarily correlated to the lambda value calculated in the residuals when that variable is correlated with an explanatory variable, we sought to also test if there were significant differences between results of correlations that we know from preliminary analyses to be sensitive to losing statistical significance depending on whether phylogenetic relatedness of taxa are being accounted for. These relationships were SD-MH and SLA-LA.

### SLA-LA Correlation

SLA has the higher lambda in general than LA so for consistency with later data we'll use SLA as the dependent variable here. The following 15 code chunks refine the tree before rerunning the same model each time.


```{r}
# Shortened redefinition of the tree being used in the initial data preparation section of this notebook
tree <- read.tree(text = pot.tree.list[[1]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

sla.la.prel1 <- pgls(sla_mm2_per_mg ~ leaf_area_cm2, data = eucs.scaled.obj, lambda = 'ML')
summary(sla.la.prel1)
```

```{r}
# now we're using tree number 2 out of the 15
tree <- read.tree(text = pot.tree.list[[2]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

sla.la.prel2 <- pgls(sla_mm2_per_mg ~ leaf_area_cm2, data = eucs.scaled.obj, lambda = 'ML')
summary(sla.la.prel2)
```

```{r}
# Shortened redefinition of the tree being used in the initial data preparation section of this notebook
tree <- read.tree(text = pot.tree.list[[3]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

sla.la.prel3 <- pgls(sla_mm2_per_mg ~ leaf_area_cm2, data = eucs.scaled.obj, lambda = 'ML')
summary(sla.la.prel3)
```

```{r}
# Shortened redefinition of the tree being used in the initial data preparation section of this notebook
tree <- read.tree(text = pot.tree.list[[4]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

sla.la.prel4 <- pgls(sla_mm2_per_mg ~ leaf_area_cm2, data = eucs.scaled.obj, lambda = 'ML')
summary(sla.la.prel4)
```

```{r}
# Shortened redefinition of the tree being used in the initial data preparation section of this notebook
tree <- read.tree(text = pot.tree.list[[5]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

sla.la.prel5 <- pgls(sla_mm2_per_mg ~ leaf_area_cm2, data = eucs.scaled.obj, lambda = 'ML')
summary(sla.la.prel5)
```

```{r}
# Shortened redefinition of the tree being used in the initial data preparation section of this notebook
tree <- read.tree(text = pot.tree.list[[6]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

sla.la.prel6 <- pgls(sla_mm2_per_mg ~ leaf_area_cm2, data = eucs.scaled.obj, lambda = 'ML')
summary(sla.la.prel6)
```

```{r}
# Shortened redefinition of the tree being used in the initial data preparation section of this notebook
tree <- read.tree(text = pot.tree.list[[7]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

sla.la.prel7 <- pgls(sla_mm2_per_mg ~ leaf_area_cm2, data = eucs.scaled.obj, lambda = 'ML')
summary(sla.la.prel7)
```

```{r}
# Shortened redefinition of the tree being used in the initial data preparation section of this notebook
tree <- read.tree(text = pot.tree.list[[8]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

sla.la.prel8 <- pgls(sla_mm2_per_mg ~ leaf_area_cm2, data = eucs.scaled.obj, lambda = 'ML')
summary(sla.la.prel8)
```

```{r}
# Shortened redefinition of the tree being used in the initial data preparation section of this notebook
tree <- read.tree(text = pot.tree.list[[9]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

sla.la.prel9 <- pgls(sla_mm2_per_mg ~ leaf_area_cm2, data = eucs.scaled.obj, lambda = 'ML')
summary(sla.la.prel9)
```

```{r}
# Shortened redefinition of the tree being used in the initial data preparation section of this notebook
tree <- read.tree(text = pot.tree.list[[10]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

sla.la.prel10 <- pgls(sla_mm2_per_mg ~ leaf_area_cm2, data = eucs.scaled.obj, lambda = 'ML')
summary(sla.la.prel10)
```

```{r}
# Shortened redefinition of the tree being used in the initial data preparation section of this notebook
tree <- read.tree(text = pot.tree.list[[11]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

sla.la.prel11 <- pgls(sla_mm2_per_mg ~ leaf_area_cm2, data = eucs.scaled.obj, lambda = 'ML')
summary(sla.la.prel11)
```

```{r}
# Shortened redefinition of the tree being used in the initial data preparation section of this notebook
tree <- read.tree(text = pot.tree.list[[12]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

sla.la.prel12 <- pgls(sla_mm2_per_mg ~ leaf_area_cm2, data = eucs.scaled.obj, lambda = 'ML')
summary(sla.la.prel12)
```

```{r}
# Shortened redefinition of the tree being used in the initial data preparation section of this notebook
tree <- read.tree(text = pot.tree.list[[13]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

sla.la.prel13 <- pgls(sla_mm2_per_mg ~ leaf_area_cm2, data = eucs.scaled.obj, lambda = 'ML')
summary(sla.la.prel13)
```

```{r}
# Shortened redefinition of the tree being used in the initial data preparation section of this notebook
tree <- read.tree(text = pot.tree.list[[14]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

sla.la.prel14 <- pgls(sla_mm2_per_mg ~ leaf_area_cm2, data = eucs.scaled.obj, lambda = 'ML')
summary(sla.la.prel14)
```

```{r}
# Shortened redefinition of the tree being used in the initial data preparation section of this notebook
tree <- read.tree(text = pot.tree.list[[15]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

sla.la.prel.null <- pgls(sla_mm2_per_mg ~ leaf_area_cm2, data = eucs.scaled.obj, lambda = 'ML')
summary(sla.la.prel.null)
```

All 15 trees produce the same non-significant relationship between SLA and LA and the lambda value was between 0.768-0.781 for all of them hence the range is 0.013 units of lambda.

### MH-SD Correlation

MH generally has the highest lambda of these two so for consistency we'll be using it as the dependent variable.


```{r}
# Shortened redefinition of the tree being used in the initial data preparation section of this notebook
tree <- read.tree(text = pot.tree.list[[1]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

mh.sd.prel1 <- pgls(max_height_m ~ stem_density_g_per_ml, data = eucs.scaled.obj, lambda = 'ML')
summary(mh.sd.prel1)
```

```{r}
# now we're using tree number 2 out of the 15
tree <- read.tree(text = pot.tree.list[[2]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

mh.sd.prel2 <- pgls(max_height_m ~ stem_density_g_per_ml, data = eucs.scaled.obj, lambda = 'ML')
summary(mh.sd.prel2)
```

```{r}
# Shortened redefinition of the tree being used in the initial data preparation section of this notebook
tree <- read.tree(text = pot.tree.list[[3]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

mh.sd.prel3 <- pgls(max_height_m ~ stem_density_g_per_ml, data = eucs.scaled.obj, lambda = 'ML')
summary(mh.sd.prel3)
```

```{r}
# Shortened redefinition of the tree being used in the initial data preparation section of this notebook
tree <- read.tree(text = pot.tree.list[[4]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

mh.sd.prel4 <- pgls(max_height_m ~ stem_density_g_per_ml, data = eucs.scaled.obj, lambda = 'ML')
summary(mh.sd.prel4)
```

```{r}
# Shortened redefinition of the tree being used in the initial data preparation section of this notebook
tree <- read.tree(text = pot.tree.list[[5]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

mh.sd.prel5 <- pgls(max_height_m ~ stem_density_g_per_ml, data = eucs.scaled.obj, lambda = 'ML')
summary(mh.sd.prel5)
```

```{r}
# Shortened redefinition of the tree being used in the initial data preparation section of this notebook
tree <- read.tree(text = pot.tree.list[[6]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

mh.sd.prel6 <- pgls(max_height_m ~ stem_density_g_per_ml, data = eucs.scaled.obj, lambda = 'ML')
summary(mh.sd.prel6)
```

```{r}
# Shortened redefinition of the tree being used in the initial data preparation section of this notebook
tree <- read.tree(text = pot.tree.list[[7]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

mh.sd.prel7 <- pgls(max_height_m ~ stem_density_g_per_ml, data = eucs.scaled.obj, lambda = 'ML')
summary(mh.sd.prel7)
```

```{r}
# Shortened redefinition of the tree being used in the initial data preparation section of this notebook
tree <- read.tree(text = pot.tree.list[[8]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

mh.sd.prel8 <- pgls(max_height_m ~ stem_density_g_per_ml, data = eucs.scaled.obj, lambda = 'ML')
summary(mh.sd.prel8)
```

```{r}
# Shortened redefinition of the tree being used in the initial data preparation section of this notebook
tree <- read.tree(text = pot.tree.list[[9]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

mh.sd.prel9 <- pgls(max_height_m ~ stem_density_g_per_ml, data = eucs.scaled.obj, lambda = 'ML')
summary(mh.sd.prel9)
```

```{r}
# Shortened redefinition of the tree being used in the initial data preparation section of this notebook
tree <- read.tree(text = pot.tree.list[[10]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

mh.sd.prel10 <- pgls(max_height_m ~ stem_density_g_per_ml, data = eucs.scaled.obj, lambda = 'ML')
summary(mh.sd.prel10)
```

```{r}
# Shortened redefinition of the tree being used in the initial data preparation section of this notebook
tree <- read.tree(text = pot.tree.list[[11]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

mh.sd.prel11 <- pgls(max_height_m ~ stem_density_g_per_ml, data = eucs.scaled.obj, lambda = 'ML')
summary(mh.sd.prel11)
```

```{r}
# Shortened redefinition of the tree being used in the initial data preparation section of this notebook
tree <- read.tree(text = pot.tree.list[[12]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

mh.sd.prel12 <- pgls(max_height_m ~ stem_density_g_per_ml, data = eucs.scaled.obj, lambda = 'ML')
summary(mh.sd.prel12)
```

```{r}
# Shortened redefinition of the tree being used in the initial data preparation section of this notebook
tree <- read.tree(text = pot.tree.list[[13]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

mh.sd.prel13 <- pgls(max_height_m ~ stem_density_g_per_ml, data = eucs.scaled.obj, lambda = 'ML')
summary(mh.sd.prel13)
```

```{r}
# Shortened redefinition of the tree being used in the initial data preparation section of this notebook
tree <- read.tree(text = pot.tree.list[[14]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

mh.sd.prel14 <- pgls(max_height_m ~ stem_density_g_per_ml, data = eucs.scaled.obj, lambda = 'ML')
summary(mh.sd.prel14)
```

```{r}
# Shortened redefinition of the tree being used in the initial data preparation section of this notebook
tree <- read.tree(text = pot.tree.list[[15]])
tree$node.label <- NULL
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

mh.sd.prel.null <- pgls(max_height_m ~ stem_density_g_per_ml, data = eucs.scaled.obj, lambda = 'ML')
summary(mh.sd.prel.null)
```

Here we can similarly see that the lambda values vary little (0.790 - 0.803) across 15 different trees and the non-significance also does not change with the p-value hovering around 0.10-0.15.


Hence, we can also conclude that the correlations are likely as unaffected by the differences in positioning of the final few taxa as the lambda values above were. Consequently, these remaining taxa can be placed as the authors see fit based on the available information and don't need to be excluded based on the uncertainty of those positions.





# Supp-Mat. Lambda Table


## Adding effects of lambda at deeper divergences i.e. outgroup taxa

Since we know that deep divergences are the ones that affect lambda the most, what if we were dealing with trees that had different or fewer outgroups (i.e. make 2 new trees, one with only Lophostemon as OG and the other using only Syncarpia as OG)

### Lophostemon as OG

```{r fig.width = 15, fig.height = 25}
tree <- read.tree(file = '../Input_data/final_tree.txt')
tree1 <- drop.tip(tree, "Syncarpia_glomulifera_subsp._glomulifera")
plot_dim1(tree1)
tree1$node.label <- NULL
tree1$tip.label <- gsub("_", " ", tree1$tip.label)
tree1 <- compute.brlen(tree1, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = tree1, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)
```


```{r}
lambda.mh.e1 <- pgls(max_height_m ~ 1, data = eucs.scaled.obj, lambda = 'ML', bounds = list(lambda = c(0.0001, 1), kappa = c(1e-6, 3), delta = c(1e-6, 3)))
summary(lambda.mh.e1)
```


```{r}
lambda.rh.e1 <- pgls(relative_height_by_girth ~ 1, data = eucs.scaled.obj, lambda = 'ML')
summary(lambda.rh.e1)
```


```{r}
lambda.std.e1 <- pgls(stem_density_g_per_ml ~ 1, data = eucs.scaled.obj, lambda = 'ML')
summary(lambda.std.e1)
```


```{r}
lambda.rbt.e1 <- pgls(relative_bt_by_girth ~ 1, data = eucs.scaled.obj, lambda = 'ML')
summary(lambda.rbt.e1)
```


```{r}
lambda.sla.e1 <- pgls(sla_mm2_per_mg ~ 1, data = eucs.scaled.obj, lambda = 'ML', bounds = list(lambda = c(0.0001, 1), kappa = c(1e-6, 3), delta = c(1e-6, 3)))
summary(lambda.sla.e1)
```


```{r}
lambda.lfa.e1 <- pgls(leaf_area_cm2 ~ 1, data = eucs.scaled.obj, lambda = 'ML')
summary(lambda.lfa.e1)
```


```{r}
lambda.lms.e1 <- pgls(leaf_mass_g ~ 1, data = eucs.scaled.obj, lambda = 'ML', bounds = list(lambda = c(0.001, 1), kappa = c(1e-6, 3), delta = c(1e-6, 3)))
summary(lambda.lms.e1)
```


```{r}
lambda.fms.e1 <- pgls(fruit_mass_mg ~ 1, data = eucs.scaled.obj, lambda = 'ML')
summary(lambda.fms.e1)
```


```{r}
lambda.fwl.e1 <- pgls(fruit_wall_width_mm ~ 1, data = eucs.scaled.obj, lambda = 'ML')
summary(lambda.fwl.e1)
```


```{r}
lambda.sms.e1 <- pgls(seed_mass_mg ~ 1, data = eucs.scaled.obj, lambda = 'ML', bounds = list(lambda = c(0.0001, 1), kappa = c(1e-6, 3), delta = c(1e-6, 3)))
summary(lambda.sms.e1)
```


```{r}
# now to recreate the table of lambda values
lambda.full.extra1 <- data.frame('Traits' = c("Max. height", "Relative Height", "Stem Density", "Relative Bark Thickness", "SLA", "Leaf Area", "Leaf Mass", "Fruit Mass", "Fruit Wall Width", "Seed Mass"), 'Lambda' =  c(lambda.mh.e1$param.CI$lambda$opt, lambda.rh.e1$param.CI$lambda$opt, lambda.std.e1$param.CI$lambda$opt, lambda.rbt.e1$param.CI$lambda$opt, lambda.sla.e1$param.CI$lambda$opt, lambda.lfa.e1$param.CI$lambda$opt, lambda.lms.e1$param.CI$lambda$opt, lambda.fms.e1$param.CI$lambda$opt, lambda.fwl.e1$param.CI$lambda$opt, lambda.sms.e1$param.CI$lambda$opt))

# extracts the two values for boundaries of the 95% confidence interval of lambda for each model
ci.val <- c(lambda.mh.e1$param.CI$lambda$ci.val, lambda.rh.e1$param.CI$lambda$ci.val, lambda.std.e1$param.CI$lambda$ci.val, lambda.rbt.e1$param.CI$lambda$ci.val, lambda.sla.e1$param.CI$lambda$ci.val, lambda.lfa.e1$param.CI$lambda$ci.val, lambda.lms.e1$param.CI$lambda$ci.val, lambda.fms.e1$param.CI$lambda$ci.val, lambda.fwl.e1$param.CI$lambda$ci.val, lambda.sms.e1$param.CI$lambda$ci.val)

# splits the CI bounds into two separate objects, one for the lower and another for the upper bound
bounds <- divide_ci(ci.val)

# constructing the full data frame of estimate and respective bounds
lambda.full.extra <- cbind(lambda.full.extra1, bounds)
## since the likely reason for the NA value next to relative height is that it's non-significance would take the lower bound below 0, which is technically below what values lambda can take (or is tested for by this function), hence for our purposes of calculating range and plotting error bars it is fine to replace it with 0 for now
lambda.full.extra[lambda.full.extra$Traits == "Relative Height", c("lower_CI")] <- 0 
## having an extra column to compare ranges
lambda.full.extra$range_CI <- lambda.full.extra$upper_CI - lambda.full.extra$lower_CI
```


Now for the other tree with Syncarpia as the outgroup.


### Syncarpia as OG


```{r fig.width = 15, fig.height = 25}
tree2 <- tree1
tree2$tip.label[tree2$tip.label == "Lophostemon confertus"] <- "Syncarpia glomulifera subsp. glomulifera"
plot_dim1(tree2)
tree2$node.label <- NULL
tree2$tip.label <- gsub("_", " ", tree2$tip.label)
tree2 <- compute.brlen(tree2, method = "Grafen", power = 1)
eucs.scaled.obj <- comparative.data(phy = tree2, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)
```


```{r}
lambda.mh.e2 <- pgls(max_height_m ~ 1, data = eucs.scaled.obj, lambda = 'ML', bounds = list(lambda = c(0.0001, 1), kappa = c(1e-6, 3), delta = c(1e-6, 3)))
summary(lambda.mh.e2)
```


```{r}
lambda.rh.e2 <- pgls(relative_height_by_girth ~ 1, data = eucs.scaled.obj, lambda = 'ML')
summary(lambda.rh.e2)
```


```{r}
lambda.std.e2 <- pgls(stem_density_g_per_ml ~ 1, data = eucs.scaled.obj, lambda = 'ML')
summary(lambda.std.e2)
```


```{r}
lambda.rbt.e2 <- pgls(relative_bt_by_girth ~ 1, data = eucs.scaled.obj, lambda = 'ML')
summary(lambda.rbt.e2)
```


```{r}
lambda.sla.e2 <- pgls(sla_mm2_per_mg ~ 1, data = eucs.scaled.obj, lambda = 'ML', bounds = list(lambda = c(0.0001, 1), kappa = c(1e-6, 3), delta = c(1e-6, 3)))
summary(lambda.sla.e2)
```


```{r}
lambda.lfa.e2 <- pgls(leaf_area_cm2 ~ 1, data = eucs.scaled.obj, lambda = 'ML')
summary(lambda.lfa.e2)
```


```{r}
lambda.lms.e2 <- pgls(leaf_mass_g ~ 1, data = eucs.scaled.obj, lambda = 'ML', bounds = list(lambda = c(0.001, 1), kappa = c(1e-6, 3), delta = c(1e-6, 3)))
summary(lambda.lms.e2)
```


```{r}
lambda.fms.e2 <- pgls(fruit_mass_mg ~ 1, data = eucs.scaled.obj, lambda = 'ML')
summary(lambda.fms.e2)
```


```{r}
lambda.fwl.e2 <- pgls(fruit_wall_width_mm ~ 1, data = eucs.scaled.obj, lambda = 'ML')
summary(lambda.fwl.e2)
```


```{r}
lambda.sms.e2 <- pgls(seed_mass_mg ~ 1, data = eucs.scaled.obj, lambda = 'ML', bounds = list(lambda = c(0.0001, 1), kappa = c(1e-6, 3), delta = c(1e-6, 3)))
summary(lambda.sms.e2)
```

```{r}
# now to recreate the table of lambda values
lambda.full.extra2 <- data.frame('Traits' = c("Max. height", "Relative Height", "Stem Density", "Relative Bark Thickness", "SLA", "Leaf Area", "Leaf Mass", "Fruit Mass", "Fruit Wall Width", "Seed Mass"), 'Lambda' =  c(lambda.mh.e2$param.CI$lambda$opt, lambda.rh.e2$param.CI$lambda$opt, lambda.std.e2$param.CI$lambda$opt, lambda.rbt.e2$param.CI$lambda$opt, lambda.sla.e2$param.CI$lambda$opt, lambda.lfa.e2$param.CI$lambda$opt, lambda.lms.e2$param.CI$lambda$opt, lambda.fms.e2$param.CI$lambda$opt, lambda.fwl.e2$param.CI$lambda$opt, lambda.sms.e2$param.CI$lambda$opt))

# extracts the two values for boundaries of the 95% confidence interval of lambda for each model
ci.val <- c(lambda.mh.e2$param.CI$lambda$ci.val, lambda.rh.e2$param.CI$lambda$ci.val, lambda.std.e2$param.CI$lambda$ci.val, lambda.rbt.e2$param.CI$lambda$ci.val, lambda.sla.e2$param.CI$lambda$ci.val, lambda.lfa.e2$param.CI$lambda$ci.val, lambda.lms.e2$param.CI$lambda$ci.val, lambda.fms.e2$param.CI$lambda$ci.val, lambda.fwl.e2$param.CI$lambda$ci.val, lambda.sms.e2$param.CI$lambda$ci.val)

# splits the CI bounds into two separate objects, one for the lower and another for the upper bound
bounds <- divide_ci(ci.val)

# constructing the full data frame of estimate and respective bounds
lambda.full.extra2 <- cbind(lambda.full.extra2, bounds)
## since the likely reason for the NA value next to relative height is that it's non-significance would take the lower bound below 0, which is technically below what values lambda can take (or is tested for by this function), hence for our purposes of calculating range and plotting error bars it is fine to replace it with 0 for now
lambda.full.extra2[lambda.full.extra2$Traits == "Relative Height", c("lower_CI")] <- 0 
## having an extra column to compare ranges
lambda.full.extra2$range_CI <- lambda.full.extra2$upper_CI - lambda.full.extra2$lower_CI
```


### Adding both trees to table

```{r}
# first to remove the summary columns at the end as we'll have to recalculate them after we add the two new columns
compare.lambda.trees2 <- compare.lambda.trees[, !names(compare.lambda.trees) %in% c("range", "min_CI_range")]

# now to make the two columns for these two extra trees that will be added to the end of the existing table from much earlier
extra.trees.lambda <- data.frame('Lambda_OG1' = lambda.full.extra$Lambda, 'Lambda_OG2' = lambda.full.extra2$Lambda)

compare.lambda.trees2 <- cbind(compare.lambda.trees2, extra.trees.lambda)

# now to recalculate the max range 
compare.lambda.trees2$range <- unlist(lapply(as.data.frame(t(as.matrix(compare.lambda.trees2[,!names(compare.lambda.trees2) %in% c('Traits')]))), function(lambdas_for_a_trait){
  range <- range(lambdas_for_a_trait)[[2]] - range(lambdas_for_a_trait)[[1]]
  return(range)
}))

# now to construct two columns with the CI ranged for each tree
extra.trees.CI.ranges <- data.frame("range_CI_OG1" = lambda.full.extra$range_CI, "range_CI_OG2" = lambda.full.extra2$range_CI)

# adding these column to the compare.range.CI.trees
compare.range.CI.trees2 <- cbind(compare.range.CI.trees, extra.trees.CI.ranges)

# now to add the minimum of this across all trees to the compare.lambda.trees2 table
compare.lambda.trees2$min_CI_range <- unlist(lapply(as.data.frame(t(as.matrix(compare.range.CI.trees2[,!names(compare.range.CI.trees2) %in% c('Traits')]))), min))
```


```{r}
# Now to add the final columns containing the values, CI bounds, and ranges for the final tree we will be using 
final.tree.lambdas <- read.csv("../Output_figures/final_tree_lambdas.csv")
final.tree.lambdas <- final.tree.lambdas[,names(final.tree.lambdas) %in% c('Traits', 'Lambda', 'lower_CI', 'upper_CI', 'range_CI')]

final.lambda.comparisons2 <- left_join(compare.lambda.trees2, final.tree.lambdas, by = 'Traits')
```


```{r}
# now to export this whole table to csv
write.csv(final.lambda.comparisons2, file = '../Output_figures/Potential_trees_and_results/suppmat_pot_tree_lambdas2.csv')
```


## Final Formatting of Table

Original horizontal version does not fit well on a page (portrait-orientation) so I'll try transposing it and see if that works.
Also, we need to round the numbers off a bit, lets to 4 dp for now.

```{r}
# first to read the previous table output back in just in case, since this is an adjustment of an existing file
final.lambda.comparisons2 <- read.csv("../Output_figures/Potential_trees_and_results/suppmat_pot_tree_lambdas2.csv")
final.lambda.comparisons.2.2 <- as.data.frame(t(final.lambda.comparisons2))

# already loaded this object but I'll rename to preserve original
df <- final.lambda.comparisons.2.2

# now to clean up unecessary rows
colnames(df) <- c("Max_height", "Relative_Height", "Stem_Density", "Relative_Bark_Thickness", "SLA", "Leaf_Area", "Leaf_Mass", "Fruit_Mass", "Fruit_Wall_Width", "Seed_Mass")
df <- df[!row.names(df) %in% c('X', 'Traits'),]

# now to round off the values
df2 <- data.frame('Max. height' = unlist(lapply(df$Max_height, num_round)),
                  'Relative Height' = unlist(lapply(df$Relative_Height, num_round)), 
                  'Stem Density' = unlist(lapply(df$Stem_Density, num_round)),
                  'Relative Bark Thickness' = unlist(lapply(df$Relative_Bark_Thickness, num_round)),
                  'SLA' = unlist(lapply(df$SLA, num_round)),
                  'Leaf Area' = unlist(lapply(df$Leaf_Area, num_round)),
                  'Leaf Mass' = unlist(lapply(df$Leaf_Mass, num_round)),
                  'Fruit Mass' = unlist(lapply(df$Fruit_Mass, num_round)),
                  'Fruit Wall Width' = unlist(lapply(df$Fruit_Wall_Width, num_round)),
                  'Seed Mass' = unlist(lapply(df$Seed_Mass, num_round)))
row.names(df2) <- row.names(df)
colnames(df2) <- colnames(df)

final.lambda.comparisons.2.3 <- df2
```

```{r}
write.csv(final.lambda.comparisons.2.3, file = '../Output_figures/Potential_trees_and_results/suppmat_pot_tree_lambdas2_vertical_rounded.csv')
```


