---
title: "Fire Correlations"
output: html_notebook
---

# Introduction

The purpose of the following notebook is to use pgls to test the nature of correlations of continuous measured traits with sprouting type (i.e. fire response) of Eucalypt taxa. This version excludes our 4 stem-only resprouting taxa from any comparisons.

# Required Packages, Functions and Data

```{r}
# Packages
library(ape)
library(caper)
library(ggplot2)
library(gridExtra)

# Functions
extract_slope <- function(model) {
  x <- summary(model)
  slope.coeff <- x$coefficients[2,1]
  return(slope.coeff)
}                   

extract_SE2 <- function(model) {
  x <- summary(model)
  slope.SE <- x$coefficients[2,2]
  return(slope.SE)
}

extract_P2 <- function(model) {
  x <- summary(model)
  P <- x$coefficients[2,c('Pr(>|t|)')]
  return(P)
}

extract_n <- function(model) {
  x <- summary(model)
  n <- x$df[[2]]
  return(n)
}

## the following is for extracting the calculated r-squared value from a pgls regression model, especially useful when list-applied to many models simultaneously
r_squared <- function(model) {
  x <- summary(model)
  x$r.squared
}

# Data Import
eucs.data <- read.csv("../Input_data/eucs_data.csv")
eucs.data <- subset(eucs.data, select = -c(X))
eucs.data[c("Sprouting_Type")] <- as.factor(eucs.data$Sprouting_Type)
eucs.data[c("Sprouting_Type")] <- as.factor(eucs.data$Sprouting_Type)

eucs.log.data <- read.csv("../Input_data/eucs_log_data.csv")
eucs.log.data <- subset(eucs.log.data, select = -c(X))
eucs.log.data[c("Sprouting_Type")] <- as.factor(eucs.log.data$Sprouting_Type)
eucs.log.data[c("Sprouting_Type2")] <- as.factor(eucs.log.data$Sprouting_Type2)

eucs.scaled.data <- read.csv("../Input_data/eucs_scaled_data.csv")
eucs.scaled.data <- subset(eucs.scaled.data, select = -c(X))
eucs.scaled.data[c("Sprouting_Type")] <- as.factor(eucs.scaled.data$Sprouting_Type)
eucs.scaled.data[c("Sprouting_Type2")] <- as.factor(eucs.scaled.data$Sprouting_Type2)
```


# Data Preparation

```{r fig.height = 30, fig.width = 15}
#input the phylogenetc hypothesis to be used
tree <- read.tree("../Input_data/final_tree.txt")
tree$tip.label <- gsub("_", " ", tree$tip.label)
eucs.test.tree <- tree 
eucs.test.tree <- compute.brlen(eucs.test.tree, method = "Grafen", power = 1) #give the tree branch lengths that
#plot_dim1(eucs.test.tree)

# now to convert all the stem-only resprouters to NA values as we discussed in both the Sprouting_Type variable and Sprouting_Type2
eucs.data$Sprouting_Type[eucs.data$Sprouting_Type3 == 'Stem_Only' & !is.na(eucs.data$Sprouting_Type)] <- NA
eucs.log.data$Sprouting_Type[eucs.log.data$Sprouting_Type3 == 'Stem_Only' & !is.na(eucs.log.data$Sprouting_Type)] <- NA
eucs.scaled.data$Sprouting_Type[eucs.scaled.data$Sprouting_Type3 == 'Stem_Only' & !is.na(eucs.scaled.data$Sprouting_Type)] <- NA

eucs.data$Sprouting_Type2[eucs.data$Sprouting_Type3 == 'Stem_Only' & !is.na(eucs.data$Sprouting_Type2)] <- NA
eucs.log.data$Sprouting_Type2[eucs.log.data$Sprouting_Type3 == 'Stem_Only' & !is.na(eucs.log.data$Sprouting_Type2)] <- NA
eucs.scaled.data$Sprouting_Type2[eucs.scaled.data$Sprouting_Type3 == 'Stem_Only' & !is.na(eucs.scaled.data$Sprouting_Type2)] <- NA


# Preparing the matching objects that allow the pgls package to link data frames to tree tips for phylogenetic correlations later
##for the un-transformed data
eucs.test <- eucs.test <- comparative.data(phy = eucs.test.tree, data = eucs.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)
eucs.test$dropped$unmatched.rows # none so all good as taxa not on tree have been removed
eucs.test$dropped$tips #none as no tips on the tree are redundant

##for the log-transformed data
eucs.log.obj <- eucs.log.obj <- comparative.data(phy = eucs.test.tree, data = eucs.log.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)
eucs.log.obj$dropped$unmatched.rows 

##for the scaled data
eucs.scaled.obj <- comparative.data(phy = eucs.test.tree, data = eucs.scaled.data, names.col = taxon, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)
eucs.scaled.obj$dropped$unmatched.rows
```


# Analyses

## OLS

### Effect of Resprouting Ability

```{r}
mh.resprout <- lm(max_height_m ~ Sprouter, data = eucs.scaled.data)
summary(mh.resprout)
```


```{r}
rh.resprout <- lm(relative_height_by_girth ~ Sprouter, data = eucs.scaled.data)
summary(rh.resprout)
```


```{r}
sd.resprout <- lm(stem_density_g_per_ml ~ Sprouter, data = eucs.scaled.data)
summary(sd.resprout)
```


```{r}
rbt.resprout <- lm(relative_bt_by_girth ~ Sprouter, data = eucs.scaled.data)
summary(rbt.resprout)
```


```{r}
sla.resprout <- lm(sla_mm2_per_mg ~ Sprouter, data = eucs.scaled.data)
summary(sla.resprout)
```


```{r}
la.resprout <- lm(leaf_area_cm2 ~ Sprouter, data = eucs.scaled.data)
summary(la.resprout)
```


```{r}
lm.resprout <- lm(leaf_mass_g ~ Sprouter, data = eucs.scaled.data)
summary(lm.resprout)
```


```{r}
fww.resprout <- lm(fruit_wall_width_mm ~ Sprouter, data = eucs.scaled.data)
summary(fww.resprout)
```


```{r}
fm.resprout <- lm(fruit_mass_mg ~ Sprouter, data = eucs.scaled.data)
summary(fm.resprout)
```


```{r}
sm.resprout <- lm(seed_mass_mg ~ Sprouter, data = eucs.scaled.data)
summary(sm.resprout)
```

Max. height and relative height are the only two traits that showed significant difference between obligate seeders and resprouters for the OLM analysis. Both were higher in obligate seeders.



### Individual Category Comparisons

#### Ref. Class Ob-Seeders

Now to run the individual comparisons between obligate seeders and either combination resprouters or lignotuber-only resprouters. We need to alter the contrasts within the factor variable to make the obligate seeder category the reference class.

```{r}
eucs.scaled.data2 <- eucs.scaled.data #creates an alternate data frame just for this analysis that has different contrasts for the three-level categorical sprouting variables
class(eucs.scaled.data$Sprouting_Type)
contrasts(eucs.scaled.data2$Sprouting_Type) <- as.matrix(data.frame('Lignotuber_Only' = c(0, 1, 0), 'Combination_Sprouter' = c(1, 0, 0), row.names = levels(eucs.scaled.data2$Sprouting_Type)))
```


```{r}
summary(lm(max_height_m ~ Sprouting_Type, data = eucs.scaled.data2))
```


```{r}
summary(lm(relative_height_by_girth ~ Sprouting_Type, data = eucs.scaled.data2))
```


```{r}
summary(lm(stem_density_g_per_ml ~ Sprouting_Type, data = eucs.scaled.data2))
```


```{r}
summary(lm(relative_bt_by_girth ~ Sprouting_Type, data = eucs.scaled.data2))
```


```{r}
summary(lm(sla_mm2_per_mg ~ Sprouting_Type, data = eucs.scaled.data2))
```


```{r}
summary(lm(leaf_area_cm2 ~ Sprouting_Type, data = eucs.scaled.data2))
```


```{r}
summary(lm(leaf_mass_g ~ Sprouting_Type, data = eucs.scaled.data2))
```


```{r}
summary(lm(fruit_wall_width_mm ~ Sprouting_Type, data = eucs.scaled.data2))
```


```{r}
summary(lm(fruit_mass_mg ~ Sprouting_Type, data = eucs.scaled.data2))
```


```{r}
summary(lm(seed_mass_mg ~ Sprouting_Type, data = eucs.scaled.data2))
```

Obligate Seeders different from basal-only resprouters by having larger MH, larger RH, smaller SD, larger RBT!, larger SLA and larger LA + (new since Pauline's draft) larger leaf mass.

Max height NON-SIGNIFICANT (but borderline) difference between obligate seeders and combination sprouters like it is in the PGLS version. 
Otherwise, only RH was significantly different between the obligate seeders and combination sprouters.


#### Ref. Class Combo sprouts

Now we're using the default contrasts that sets combination resprouters as the reference category (alphabetical order).

```{r}
contrasts(eucs.scaled.data$Sprouting_Type)
summary(lm(max_height_m ~ Sprouting_Type, data = eucs.scaled.data))
```


```{r}
summary(lm(max_height_m ~ Sprouting_Type, data = eucs.scaled.data))
```
Weakened in PGLS -1.5 and r2 = 0.40

```{r}
summary(lm(relative_height_by_girth ~ Sprouting_Type, data = eucs.scaled.data))
```

Practically the same in PGLS

```{r}
summary(lm(stem_density_g_per_ml ~ Sprouting_Type, data = eucs.scaled.data))
```

Slightly weakened in PGLS 0.65 and r2 = 0.068

```{r}
summary(lm(relative_bt_by_girth ~ Sprouting_Type, data = eucs.scaled.data))
```

Practically the same in PGLS

```{r}
summary(lm(sla_mm2_per_mg ~ Sprouting_Type, data = eucs.scaled.data))
```

Weakened significantly by PGLS down to -0.91 and r2 = 0.13


```{r}
summary(lm(leaf_area_cm2 ~ Sprouting_Type, data = eucs.scaled.data))
```

 Weakened in PGLS -1.15 and r2 = 0.20

```{r}
summary(lm(leaf_mass_g ~ Sprouting_Type, data = eucs.scaled.data))
```

Practically the same in PGLS

```{r}
summary(lm(fruit_wall_width_mm ~ Sprouting_Type, data = eucs.scaled.data))
```

Slightly weakened in PGLS 0.53 and r2 = 0.041

```{r}
summary(lm(fruit_mass_mg ~ Sprouting_Type, data = eucs.scaled.data))
```

Weakened significantly by PGLS 0.13 and r2 = 0.010


```{r}
summary(lm(seed_mass_mg ~ Sprouting_Type, data = eucs.scaled.data))
```

Weakened significantly to non-significance by PGLS 0.12 and r2 = 0.008

Basal-only resprouters are shorter, stouter, have denser wood, thinner bark, lower-sla smaller and lighter leaves, thicker fruit walls AS WELL as HEAVIER FRUITS and SEEDS than combination sprouters.
Now heavier fruits no longer significant

## PGLS

### Effect of Resprouting Ability

```{r}
max.height.fire <- pgls(max_height_m ~ Sprouter, data = eucs.scaled.obj, lambda = "ML")
summary(max.height.fire)
```


```{r}
rel.height.fire <- pgls(relative_height_by_girth ~ Sprouter, data = eucs.scaled.obj, lambda = 'ML')
summary(rel.height.fire)
```


```{r}
stem.density.fire <- pgls(stem_density_g_per_ml ~ Sprouter, data = eucs.scaled.obj, lambda = 'ML')
summary(stem.density.fire)
#plot(stem_density_g_per_ml ~ Sprouter, data = eucs.scaled.data)
```


```{r}
bark.fire <- pgls(relative_bt_by_girth ~ Sprouter, data = eucs.scaled.obj, lambda = 'ML')
summary(bark.fire)
```


```{r}
sla.fire <- pgls(sla_mm2_per_mg ~ Sprouter, data = eucs.scaled.obj, lambda = 'ML')
summary(sla.fire)
```


```{r}
leaf.area.fire <- pgls(leaf_area_cm2 ~ Sprouter, data = eucs.scaled.obj, lambda = 'ML')
summary(leaf.area.fire)
```


```{r}
leaf.mass.fire <- pgls(leaf_mass_g ~ Sprouter, data = eucs.scaled.obj, lambda = 'ML')
summary(leaf.mass.fire)
```


```{r}
fruit.wall.fire <- pgls(fruit_wall_width_mm ~ Sprouter, data = eucs.scaled.obj, lambda = 'ML')
summary(fruit.wall.fire)
plot(fruit_wall_width_mm ~ Sprouter, data = eucs.log.data)
```


```{r}
fruit.mass.fire <- pgls(fruit_mass_mg ~ Sprouter, data = eucs.scaled.obj, lambda = 'ML',  bounds = list(lambda = c(0.001, 1), kappa = c(1e-6, 3), delta = c(1e-6, 3)))
summary(fruit.mass.fire)
```


```{r}
seed.mass.fire <- pgls(seed_mass_mg ~ Sprouter, data = eucs.scaled.obj, lambda = 'ML')
summary(seed.mass.fire)
```

Only Max. Height and Relative Height were significantly different between obligate seeders and sprouters. Both were higher in obligate seeders.

### Individual Category Comparisons

#### Ref. Class Ob-seeders

First need to set the contrasts within the Sprouting Type variable to treat Obligate seeders as the reference category.

```{r}
eucs.scaled.obj2 <- eucs.scaled.obj
contrasts(eucs.scaled.obj2$data$Sprouting_Type) <- as.matrix(data.frame('Lignotuber_Only' = c(0, 1, 0), 'Combination_Sprouter' = c(1, 0, 0), row.names = row.names(contrasts(eucs.scaled.obj2$data$Sprouting_Type))))
```


```{r}
summary(pgls(max_height_m ~ Sprouting_Type, data = eucs.scaled.obj2, lambda = 'ML'))
# summary(pgls(max_height_m ~ Sprouting_Type, data = eucs.scaled.obj.con, lambda = 'ML'))
# contrasts(eucs.scaled.obj.con$data$Sprouting_Type)
```


```{r}
summary(pgls(relative_height_by_girth ~ Sprouting_Type, data = eucs.scaled.obj2, lambda = 'ML'))
```

```{r}
summary(pgls(stem_density_g_per_ml ~ Sprouting_Type, data = eucs.scaled.obj2, lambda = 'ML'))
```

```{r}
summary(pgls(relative_bt_by_girth ~ Sprouting_Type, data = eucs.scaled.obj2, lambda = 'ML'))
```

```{r}
summary(pgls(sla_mm2_per_mg ~ Sprouting_Type, data = eucs.scaled.obj2, lambda = 'ML'))

```

```{r}
summary(pgls(leaf_area_cm2 ~ Sprouting_Type, data = eucs.scaled.obj2, lambda = 'ML'))
```

```{r}
summary(pgls(leaf_mass_g ~ Sprouting_Type, data = eucs.scaled.obj2, lambda = 'ML'))
```

```{r}
summary(pgls(fruit_wall_width_mm ~ Sprouting_Type, data = eucs.scaled.obj2, lambda = 'ML'))
```


```{r}
summary(pgls(fruit_mass_mg ~ Sprouting_Type, data = eucs.scaled.obj2, lambda = 'ML'))
```


```{r}
summary(pgls(seed_mass_mg ~ Sprouting_Type, data = eucs.scaled.obj2, lambda = 'ML'))
```

Max. Height (and now RH) are the only traits that is different between obligate seeders and combination resprouters, MH was not significant in the OLS analysis.
MH shorter, RH stouter, SD denser, RBT thinner, SLA smaller,  LA smaller, and now LM lighter were different for basal-only resprouters compared to ob-seeders. Note FWW was not different, though this is the same as the OLM analysis.


#### Ref. Class Combo Sprouts


The following method does not employ Helmert contrasts and the like but simply uses the default contrasts when the three categories are treated as three levels of a single categorical variable, which sets Combination_Sprouter as the reference class and the others are represented by their difference to it. This allows the measurement of the difference between the obligate seeders and combination sprouters as well.

```{r}
# quich verification that contrasts place combination resprouters as the reference category
contrasts(eucs.scaled.obj$data$Sprouting_Type)

# start the correlations
phy.comboref.mh <-pgls(max_height_m ~ Sprouting_Type, data = eucs.scaled.obj, lambda = 'ML')
summary(phy.comboref.mh)
```


```{r}
phy.comboref.rh <- pgls(relative_height_by_girth ~ Sprouting_Type, data = eucs.scaled.obj, lambda = 'ML')
summary(phy.comboref.rh)
```

```{r}
phy.comboref.sd <- pgls(stem_density_g_per_ml ~ Sprouting_Type, data = eucs.scaled.obj, lambda = 'ML')
summary(phy.comboref.sd)
```

```{r}
phy.comboref.rbt <- pgls(relative_bt_by_girth ~ Sprouting_Type, data = eucs.scaled.obj, lambda = 'ML')
summary(phy.comboref.rbt)
```

```{r}
phy.comboref.sla <- pgls(sla_mm2_per_mg ~ Sprouting_Type, data = eucs.scaled.obj, lambda = 'ML')
summary(phy.comboref.sla)
```

```{r}
phy.comboref.la <- pgls(leaf_area_cm2 ~ Sprouting_Type, data = eucs.scaled.obj, lambda = 'ML')
summary(phy.comboref.la)
```

```{r}
phy.comboref.lm <- pgls(leaf_mass_g ~ Sprouting_Type, data = eucs.scaled.obj, lambda = 'ML')
summary(phy.comboref.lm)
```

```{r}
phy.comboref.fww <- pgls(fruit_wall_width_mm ~ Sprouting_Type, data = eucs.scaled.obj, lambda = 'ML')
summary(phy.comboref.fww)
```

```{r}
phy.comboref.fm <- pgls(fruit_mass_mg ~ Sprouting_Type, data = eucs.scaled.obj, lambda = 'ML')
summary(phy.comboref.fm)
```

```{r}
phy.comboref.sm <- pgls(seed_mass_mg ~ Sprouting_Type, data = eucs.scaled.obj, lambda = 'ML')
summary(phy.comboref.sm)
```

Differences between combinations resprouters and basal-only resprouters were very similar with the addition of FWW (thicker) also being additionally different to when basal-repsrouters were compared to obligate seeders instead.
Compared to the OLS analysis, the significantly heavier seeds and fruits of lignotuber-only resprouters vs lighter in combination resprouters was no longer significant and the heavier FWW of ligno-sprouters was weakened until only borderline significant.

## Checking variably sprouting taxa

First need to check the contrasts within the Sprouting Type variable for the comparison where combination resprouters are the reference category as that had the most interesting information.

```{r}
# the following should show a 0 0 for the row for combination resprouters
contrasts(eucs.scaled.obj$data$Sprouting_Type2)
```


```{r}
variables.test.mh <- pgls(max_height_m ~ Sprouting_Type2, data = eucs.scaled.obj, lambda = 'ML')
summary(variables.test.mh)
# summary(pgls(max_height_m ~ Sprouting_Type, data = eucs.scaled.obj.con, lambda = 'ML'))
# contrasts(eucs.scaled.obj.con$data$Sprouting_Type)
```


```{r}
variables.test.rh <- pgls(relative_height_by_girth ~ Sprouting_Type2, data = eucs.scaled.obj, lambda = 'ML')
summary(variables.test.rh)
```

```{r}
variables.test.sd <- pgls(stem_density_g_per_ml ~ Sprouting_Type2, data = eucs.scaled.obj, lambda = 'ML')
summary(variables.test.sd)
```

```{r}
variables.test.rbt <- pgls(relative_bt_by_girth ~ Sprouting_Type2, data = eucs.scaled.obj, lambda = 'ML')
summary(variables.test.rbt)
```

```{r}
variables.test.sla <- pgls(sla_mm2_per_mg ~ Sprouting_Type2, data = eucs.scaled.obj, lambda = 'ML')
summary(variables.test.sla)

```

```{r}
variables.test.la <- pgls(leaf_area_cm2 ~ Sprouting_Type2, data = eucs.scaled.obj, lambda = 'ML')
summary(variables.test.la)
```

```{r}
variables.test.lm <- pgls(leaf_mass_g ~ Sprouting_Type2, data = eucs.scaled.obj, lambda = 'ML')
summary(variables.test.lm)
```

```{r}
variables.test.fww <- pgls(fruit_wall_width_mm ~ Sprouting_Type2, data = eucs.scaled.obj, lambda = 'ML')
summary(variables.test.fww)
```


```{r}
variables.test.fm <- pgls(fruit_mass_mg ~ Sprouting_Type2, data = eucs.scaled.obj, lambda = 'ML')
summary(variables.test.fm)
```


```{r}
variables.test.sm <- pgls(seed_mass_mg ~ Sprouting_Type2, data = eucs.scaled.obj, lambda = 'ML')
summary(variables.test.sm)
```

Max. Height is the only trait that is different between obligate seeders and combination resprouters, this was not true for the OLM analysis.
MH shorter, RH stouter, SD denser, RBT thinner, SLA smaller and LA smaller were different between obligate seeders and basal-only resprouters. Note LM and FWW were not different, though this is the same as the OLM analysis.

### Extracting data

```{r}
variables.tests <- list(variables.test.mh, variables.test.rh, variables.test.sd, variables.test.rbt, variables.test.sla, variables.test.la, variables.test.lm, variables.test.fww, variables.test.fm, variables.test.sm)


variables.coeff1 <- data.frame('Effect_Size' = unlist(lapply(variables.tests, extract_slope)), 'S.E.' = unlist(lapply(variables.tests, extract_SE2)) )

class(extract_n(variables.test.mh))
summary(variables.test.mh)
full.coeff.excl.vars <- data.frame('Trait' = c('MH', 'RH', 'SD', 'RBT', 'SLA', 'LA', 'LM', 'FWW', 'FM', 'SM'), 'Slope' = paste(variables.coeff1$Effect_Size, variables.coeff1$S.E., sep=" ± "), "R2" = unlist(lapply(variables.tests, r_squared)), "n" = unlist(lapply(variables.tests, extract_n)) + 2, "P_value" = unlist(lapply(variables.tests, extract_P2)))

# combined.coeff.excl.vars <- data.frame('Variable_Taxa_Excluded' = lapply(paste(full.coeff.excl.vars$Slope, full.coeff.excl.vars$R2, full.coeff.excl.vars$n, full.coeff.excl.vars$P_value, sep = "\n"), cat))
# 
# cat(paste(full.coeff.excl.vars$n, full.coeff.excl.vars$R2, sep = "\n"))

# NEXT WE NEED TO DO THE SAME WITH THE NORMAL PGLS THAT INCLUDES THE VARIABLE TAXA WITHIN COMBO SPROUTERS AND THEN ADD THE R-VALUE AND MAYBE P-VALES INTO THE SAME CELL VIA THE "\n" IN THE PASTE FUNCTION

pgls.comboref.models <- list(phy.comboref.mh, phy.comboref.rh, phy.comboref.sd, phy.comboref.rbt, phy.comboref.sla, phy.comboref.la, phy.comboref.lm, phy.comboref.fww, phy.comboref.fm, phy.comboref.sm)

pgls.comboref.coeff <- data.frame('Effect_Size_incl' = unlist(lapply(pgls.comboref.models, extract_slope)), 'S.E.incl' = unlist(lapply(pgls.comboref.models, extract_SE2)))

full.coeff.incl.vars <- data.frame("Traits" = c('MH', 'RH', 'SD', 'RBT', 'SLA', 'LA', 'LM', 'FWW', 'FM', 'SM'), 'Slope' = paste(pgls.comboref.coeff$Effect_Size_incl, pgls.comboref.coeff$S.E.incl, sep=" ± "), "R2" = unlist(lapply(pgls.comboref.models, r_squared)), "n" = unlist(lapply(pgls.comboref.models, extract_n)) + 2, "P_value" = unlist(lapply(pgls.comboref.models, extract_P2)))

write.csv(full.coeff.incl.vars, file = "../Output_figures/incl_variable_taxa.csv")
write.csv(full.coeff.excl.vars, file = "../Output_figures/excl_variable_taxa.csv")
# write.table(full.coeff.incl.vars, file = "Output_figures/incl_variable_taxa", eol = ',', sep = ',')
# write.table(t(full.coeff.excl.vars), file = "Output_figures/excl_variable_taxa", eol = ',', row.names = FALSE)

# A = 'hi\nand'
# cat(A)
```

R2 values comparison
    OLS     PGLS
MH  0.56		0.40
RH  0.10		0.10
SD  0.15		0.07
RBT 0.22		0.22
SLA 0.37		0.13
LA  0.36		0.20
LM  0.11		0.11
FWW 0.11		0.04
FM  0.03		0.01
SM  0.04		0.01

# Plots

The following is an exploration of the different possible methods of visualising the above results. For the final version used in the paper skip the the relevant subheadings within the heading "Without the NA column".

## Just Dots

```{r}
ggplot(data = eucs.scaled.data, aes(x = Sprouting_Type, y = max_height_m, fill = Sprouting_Type)) + geom_point(aes(y = max_height_m, color = Sprouting_Type), position = position_jitter(width = .15, seed =1), size = 1.5, alpha = 0.9) + scale_x_discrete(limits = c(NA, "Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only")) + coord_flip() 

ggplot(data = eucs.scaled.data, aes(x = Sprouting_Type, y = relative_height_by_girth, fill = Sprouting_Type)) + geom_point(aes(y = relative_height_by_girth, color = Sprouting_Type), position = position_jitter(width = .15, seed = 1), size = 1.5, alpha = 0.8) + scale_x_discrete(limits = c(NA, "Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only")) + coord_flip()

ggplot(data = eucs.scaled.data, aes(x = Sprouting_Type, y = stem_density_g_per_ml, fill = Sprouting_Type)) + geom_point(aes(y = stem_density_g_per_ml, color = Sprouting_Type), position = position_jitter(width = .15, seed = 1), size = 1.5, alpha = 0.8) + scale_x_discrete(limits = c(NA, "Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only")) + coord_flip()

ggplot(data = eucs.scaled.data, aes(x = Sprouting_Type, y = relative_bt_by_girth, fill = Sprouting_Type)) + geom_point(aes(y = relative_bt_by_girth, color = Sprouting_Type), position = position_jitter(width = .15, seed = 1), size = 1.5, alpha = 0.8) + scale_x_discrete(limits = c(NA, "Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only")) + coord_flip()

ggplot(data = eucs.scaled.data, aes(x = Sprouting_Type, y = sla_mm2_per_mg, fill = Sprouting_Type)) + geom_point(aes(y = sla_mm2_per_mg, color = Sprouting_Type), position = position_jitter(width = .15), size = 1.5, alpha = 0.8) + scale_x_discrete(limits = c(NA, "Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only")) + coord_flip()

ggplot(data = eucs.scaled.data, aes(x = Sprouting_Type, y = leaf_area_cm2, fill = Sprouting_Type)) + geom_point(aes(y = leaf_area_cm2, color = Sprouting_Type), position = position_jitter(width = .15), size = 1.5, alpha = 0.8) + scale_x_discrete(limits = c(NA, "Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only")) + coord_flip()

ggplot(data = eucs.scaled.data, aes(x = Sprouting_Type, y = leaf_mass_g, fill = Sprouting_Type)) + geom_point(aes(y = leaf_mass_g, color = Sprouting_Type), position = position_jitter(width = .15), size = 1.5, alpha = 0.8) + scale_x_discrete(limits = c(NA, "Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only")) + coord_flip()

ggplot(data = eucs.scaled.data, aes(x = Sprouting_Type, y = fruit_wall_width_mm, fill = Sprouting_Type)) + geom_point(aes(y = fruit_wall_width_mm, color = Sprouting_Type), position = position_jitter(width = .15), size = 1.5, alpha = 0.8) + scale_x_discrete(limits = c(NA, "Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only")) + coord_flip()

ggplot(data = eucs.scaled.data, aes(x = Sprouting_Type, y = fruit_mass_mg, fill = Sprouting_Type)) + geom_point(aes(y = fruit_mass_mg, color = Sprouting_Type), position = position_jitter(width = .15), size = 1.5, alpha = 0.8) + scale_x_discrete(limits = c(NA, "Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only")) + coord_flip()

ggplot(data = eucs.scaled.data, aes(x = Sprouting_Type, y = seed_mass_mg, fill = Sprouting_Type)) + geom_point(aes(y = seed_mass_mg, color = Sprouting_Type), position = position_jitter(width = .15), size = 1.5, alpha = 0.8) + scale_x_discrete(limits = c(NA, "Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only")) + coord_flip()

# beeswarm(fruit_mass_mg ~ Regen..Strategy, data = boxplot.log.data, horizontal = TRUE, col = c("Green", "Blue", "Orange"))
# beeswarm(seed_mass_mg ~ Regen..Strategy, data = boxplot.log.data, horizontal = TRUE, col = c("Green", "Blue", "Orange"))
# 
# ggplot(data = eucs.scaled.data, aes(x = Sprouting_Type, y = fruit_mass_mg, fill = Sprouting_Type)) + geom_point(aes(y = sla_mm2_per_mg, color = Sprouting_Type), size = .5, alpha = 0.8) + scale_x_discrete(limits = c(NA, "Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only")) + coord_flip()
# 
# ggplot(data = eucs.scaled.data, aes(x = Sprouting_Type, y = seed_mass_mg, fill = Sprouting_Type)) + geom_point(aes(y = sla_mm2_per_mg, color = Sprouting_Type), position = position_jitter(width = 0.1, seed = 1), size = .5, alpha = 0.8) + scale_x_discrete(limits = c(NA, "Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only")) + coord_flip()
# 
# stripchart(eucs.scaled.data$seed_mass_mg ~ eucs.scaled.data$Sprouting_Type, col = c("Orange", "Green", "Blue"))
# 
# ggplot(eucs.scaled.data, aes(x=Sprouting_Type, y=seed_mass_mg)) + geom_point(size = 3) + scale_x_discrete(limits = c(NA, "Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only")) + coord_flip()
#   geom_jitter(position=position_jitter(0.2))

```
The jitter gives the data a vertical dimension at a randomo height such that we can see the data a lot more clearly and where it is more or less dense.

But it ay be better to include the boxplot to this data too or simply just the whole raincloud plot.


## Dot-and-box Plots

Now just with the datapoints along with the boxplot and no density violin plot

```{r}
pdf("../Output_figures/boxplot_fire_response_vs_traits.pdf")
ggplot(data = eucs.scaled.data, 
         aes(x = Sprouting_Type, y = max_height_m, fill = Sprouting_Type)) +
  geom_point(aes(y = max_height_m, color = Sprouting_Type), position = position_jitter(width = .15, seed =1), size = 1.5, alpha = 0.6) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 5.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + scale_x_discrete(limits = c(NA, "Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only")) + labs(y="Maximum Height", x = "Post-Fire Regeneration Strategy") +
  coord_flip()

ggplot(data = eucs.scaled.data, 
         aes(x = Sprouting_Type, y = relative_height_by_girth, fill = Sprouting_Type)) +
  geom_point(aes(y = relative_height_by_girth, color = Sprouting_Type), position = position_jitter(width = .15, seed =1), size = 1.5, alpha = 0.6) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 5.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + scale_x_discrete(limits = c(NA, "Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only")) + labs(y="Relative Height by Girth", x = "Post-Fire Regeneration Strategy") +
  coord_flip() 

ggplot(data = eucs.scaled.data, 
         aes(x = Sprouting_Type, y = stem_density_g_per_ml, fill = Sprouting_Type)) +
  geom_point(aes(y = stem_density_g_per_ml, color = Sprouting_Type), position = position_jitter(width = .15, seed =1), size = 1.5, alpha = 0.6) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 5.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + scale_x_discrete(limits = c(NA, "Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only")) + labs(y="Stem Density", x = "Post-Fire Regeneration Strategy") +
  coord_flip() 

ggplot(data = eucs.scaled.data, 
         aes(x = Sprouting_Type, y = relative_bt_by_girth, fill = Sprouting_Type)) +
  geom_point(aes(y = relative_bt_by_girth, color = Sprouting_Type), position = position_jitter(width = .15, seed =1), size = 1.5, alpha = 0.6) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 5.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + scale_x_discrete(limits = c(NA, "Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only")) + labs(y="Relative Bark Thickness by Girth", x = "Post-Fire Regeneration Strategy") +
  coord_flip() 

ggplot(data = eucs.scaled.data, 
         aes(x = Sprouting_Type, y = sla_mm2_per_mg, fill = Sprouting_Type)) +
  geom_point(aes(y = sla_mm2_per_mg, color = Sprouting_Type), position = position_jitter(width = .15, seed =1), size = 1.5, alpha = 0.6) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 5.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + scale_x_discrete(limits = c(NA, "Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only")) + labs(y="SLA", x = "Post-Fire Regeneration Strategy") +
  coord_flip() 

ggplot(data = eucs.scaled.data, 
         aes(x = Sprouting_Type, y = leaf_area_cm2, fill = Sprouting_Type)) +
  geom_point(aes(y = leaf_area_cm2, color = Sprouting_Type), position = position_jitter(width = .15, seed =1), size = 1.5, alpha = 0.6) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 5.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + scale_x_discrete(limits = c(NA, "Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only")) + labs(y="Leaf Area", x = "Post-Fire Regeneration Strategy") +
  coord_flip() 

ggplot(data = eucs.scaled.data, 
         aes(x = Sprouting_Type, y = leaf_mass_g, fill = Sprouting_Type)) +
  geom_point(aes(y = leaf_mass_g, color = Sprouting_Type), position = position_jitter(width = .15, seed =1), size = 1.5, alpha = 0.6) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 5.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + scale_x_discrete(limits = c(NA, "Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only")) + labs(y="Leaf Mass", x = "Post-Fire Regeneration Strategy") +
  coord_flip() 

ggplot(data = eucs.scaled.data, 
         aes(x = Sprouting_Type, y = fruit_wall_width_mm, fill = Sprouting_Type)) +
  geom_point(aes(y = fruit_wall_width_mm, color = Sprouting_Type), position = position_jitter(width = .15, seed =1), size = 1.5, alpha = 0.6) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 5.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + scale_x_discrete(limits = c(NA, "Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only")) + labs(y="Fruit Wall Width", x = "Post-Fire Regeneration Strategy") +
  coord_flip() 
  
ggplot(data = eucs.scaled.data, 
         aes(x = Sprouting_Type, y = fruit_mass_mg, fill = Sprouting_Type)) +
  geom_point(aes(y = fruit_mass_mg, color = Sprouting_Type), position = position_jitter(width = .15, seed =1), size = 1.5, alpha = 0.6) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 5.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + scale_x_discrete(limits = c(NA, "Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only")) + labs(y="Fruit Mass", x = "Post-Fire Regeneration Strategy") +
  coord_flip() 

ggplot(data = eucs.scaled.data, 
         aes(x = Sprouting_Type, y = seed_mass_mg, fill = Sprouting_Type)) +
  geom_point(aes(y = seed_mass_mg, color = Sprouting_Type), position = position_jitter(width = .15, seed =1), size = 1.5, alpha = 0.6) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 5.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + scale_x_discrete(limits = c(NA, "Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only")) + labs(y="Seed Mass", x = "Post-Fire Regeneration Strategy") +
  coord_flip() 
dev.off()
```

## Without the NA column

### For looking at all
```{r}
pdf("../Output_figures/boxplot_fire_response_vs_traits.pdf")
ggplot(data = eucs.data, 
         aes(x = Sprouting_Type, y = max_height_m, fill = Sprouting_Type)) +
  geom_point(aes(y = max_height_m, color = Sprouting_Type), position = position_jitter(width = .15, seed =1), size = 1.5, alpha = 0.6) + scale_color_manual(values = c("Combination_Sprouter" = "blue",
                                 "Lignotuber_Only"="magenta",
                                 "Obligate_Seeder"="brown")) + 
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 3.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + 
  scale_fill_manual(breaks = eucs.data$Sprouting_Type, 
                    values = c("Combination_Sprouter" = "blue",
                                "Lignotuber_Only"="magenta",
                                "Obligate_Seeder"="brown")) + scale_x_discrete(limits = c("Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only"), labels = c("Combination Resprouter", "Obligate Seeder", "Basal-Only Resprouter")) + labs(y="Maximum Height (m)", x = "Post-Fire Regeneration Strategy") + scale_y_continuous(trans = "log10", breaks = c(1, 3, 10, 30, 100)) +coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))

ggplot(data = eucs.data, 
         aes(x = Sprouting_Type, y = relative_height_by_girth, fill = Sprouting_Type)) +
  geom_point(aes(y = relative_height_by_girth, color = Sprouting_Type), position = position_jitter(width = .15, seed =1), size = 1.5, alpha = 0.6) + scale_color_manual(values = c("Combination_Sprouter" = "blue",
                                 "Lignotuber_Only"="magenta",
                                 "Obligate_Seeder"="brown")) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 3.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + 
  scale_fill_manual(breaks = eucs.data$Sprouting_Type, 
                    values = c("Combination_Sprouter" = "blue",
                                "Lignotuber_Only"="magenta",
                                "Obligate_Seeder"="brown")) + scale_x_discrete(limits = c("Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only"), labels = c("Combination Resprouter", "Obligate Seeder", "Basal-Only Resprouter")) + labs(y="Relative Height by Girth (m)", x = "Post-Fire Regeneration Strategy") + scale_y_continuous(trans = "log10", breaks = c(6, 12, 24)) + 
  coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))

ggplot(data = eucs.data, 
         aes(x = Sprouting_Type, y = stem_density_g_per_ml, fill = Sprouting_Type)) +
  geom_point(aes(y = stem_density_g_per_ml, color = Sprouting_Type), position = position_jitter(width = .15, seed =1), size = 1.5, alpha = 0.6) + scale_color_manual(values = c("Combination_Sprouter" = "blue",
                                 "Lignotuber_Only"="magenta",
                                 "Obligate_Seeder"="brown")) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 3.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + 
  scale_fill_manual(breaks = eucs.data$Sprouting_Type, 
                    values = c("Combination_Sprouter" = "blue",
                                "Lignotuber_Only"="magenta",
                                "Obligate_Seeder"="brown")) + scale_x_discrete(limits = c("Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only"), labels = c("Combination Resprouter", "Obligate Seeder", "Basal-Only Resprouter")) + labs(y="Stem Density (g/mL)", x = "Post-Fire Regeneration Strategy") + scale_y_continuous(trans = "log10", breaks = c(0.55, 0.65, 0.75, 0.85)) +
  coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) 

ggplot(data = eucs.data, 
         aes(x = Sprouting_Type, y = relative_bt_by_girth, fill = Sprouting_Type)) +
  geom_point(aes(y = relative_bt_by_girth, color = Sprouting_Type), position = position_jitter(width = .15, seed =1), size = 1.5, alpha = 0.6) + scale_color_manual(values = c("Combination_Sprouter" = "blue",
                                 "Lignotuber_Only"="magenta",
                                 "Obligate_Seeder"="brown")) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 3.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + 
  scale_fill_manual(breaks = eucs.data$Sprouting_Type, 
                    values = c("Combination_Sprouter" = "blue",
                                "Lignotuber_Only"="magenta",
                                "Obligate_Seeder"="brown")) + scale_x_discrete(limits = c("Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only"), labels = c("Combination Resprouter", "Obligate Seeder", "Basal-Only Resprouter")) + labs(y="Relative Bark Thickness by Girth", x = "Post-Fire Regeneration Strategy") + scale_y_continuous(trans = "log10", breaks = c(5, 10, 20, 40)) +
  coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) 


ggplot(data = eucs.data, 
         aes(x = Sprouting_Type, y = sla_mm2_per_mg, fill = Sprouting_Type)) + geom_point(aes(y = sla_mm2_per_mg, color = Sprouting_Type), position = position_jitter(width = .15, seed =1), size = 1.5, alpha = 0.6) + scale_color_manual(values = c("Combination_Sprouter" = "blue",
                                 "Lignotuber_Only"="magenta",
                                 "Obligate_Seeder"="brown")) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 3.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + 
  scale_fill_manual(breaks = eucs.data$Sprouting_Type, 
                    values = c("Combination_Sprouter" = "blue",
                                "Lignotuber_Only"="magenta",
                                "Obligate_Seeder"="brown")) + scale_x_discrete(limits = c("Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only"), labels = c("Combination Resprouter", "Obligate Seeder", "Basal-Only Resprouter")) + labs(y = "SLA (mm2/mg)", x = "Post-Fire Regeneration Strategy") + scale_y_continuous(trans = "log10", breaks = c(2, 4, 8)) + 
    coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))


  
ggplot(data = eucs.data, 
         aes(x = Sprouting_Type, y = leaf_area_cm2, fill = Sprouting_Type)) +
  geom_point(aes(y = leaf_area_cm2, color = Sprouting_Type), position = position_jitter(width = .15, seed =1), size = 1.5, alpha = 0.6) + scale_color_manual(values = c("Combination_Sprouter" = "blue",
                                 "Lignotuber_Only"="magenta",
                                 "Obligate_Seeder"="brown")) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 3.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + 
  scale_fill_manual(breaks = eucs.data$Sprouting_Type, 
                    values = c("Combination_Sprouter" = "blue",
                                "Lignotuber_Only"="magenta",
                                "Obligate_Seeder"="brown")) + scale_x_discrete(limits = c("Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only")) + labs(y="Leaf Area (cm2)", x = "Post-Fire Regeneration Strategy") + scale_y_continuous(trans = "log10", breaks = c(3, 10, 30, 100)) +
  coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))

ggplot(data = eucs.data, 
         aes(x = Sprouting_Type, y = leaf_mass_g, fill = Sprouting_Type)) +
  geom_point(aes(y = leaf_mass_g, color = Sprouting_Type), position = position_jitter(width = .15, seed =1), size = 1.5, alpha = 0.6) + scale_color_manual(values = c("Combination_Sprouter" = "blue",
                                 "Lignotuber_Only"="magenta",
                                 "Obligate_Seeder"="brown")) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 3.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + 
  scale_fill_manual(breaks = eucs.data$Sprouting_Type, 
                    values = c("Combination_Sprouter" = "blue",
                                "Lignotuber_Only"="magenta",
                                "Obligate_Seeder"="brown")) + scale_x_discrete(limits = c("Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only"), labels = c("Combination Resprouter", "Obligate Seeder", "Basal-Only Resprouter")) + labs(y="Leaf Mass (g)", x = "Post-Fire Regeneration Strategy") + scale_y_continuous(trans = "log10", breaks = c(0.1, 0.5, 2.5)) +
  coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) 

ggplot(data = eucs.data, 
         aes(x = Sprouting_Type, y = fruit_wall_width_mm, fill = Sprouting_Type)) +
  geom_point(aes(y = fruit_wall_width_mm, color = Sprouting_Type), position = position_jitter(width = .15, seed =1), size = 1.5, alpha = 0.6) + scale_color_manual(values = c("Combination_Sprouter" = "blue",
                                 "Lignotuber_Only"="magenta",
                                 "Obligate_Seeder"="brown")) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 3.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + 
  scale_fill_manual(breaks = eucs.data$Sprouting_Type, 
                    values = c("Combination_Sprouter" = "blue",
                                "Lignotuber_Only"="magenta",
                                "Obligate_Seeder"="brown")) + scale_x_discrete(limits = c("Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only"), labels = c("Combination Resprouter", "Obligate Seeder", "Basal-Only Resprouter")) + labs(y="Fruit Wall Width (mm)", x = "Post-Fire Regeneration Strategy") + scale_y_continuous(trans = "log10", breaks = c(0.5, 1, 2, 3, 4)) +
  coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) 
  
ggplot(data = eucs.data, 
         aes(x = Sprouting_Type, y = fruit_mass_mg, fill = Sprouting_Type)) +
  geom_point(aes(y = fruit_mass_mg, color = Sprouting_Type), position = position_jitter(width = .15, seed =1), size = 1.5, alpha = 0.6) + scale_color_manual(values = c("Combination_Sprouter" = "blue",
                                 "Lignotuber_Only"="magenta",
                                 "Obligate_Seeder"="brown")) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 3.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + 
  scale_fill_manual(breaks = eucs.data$Sprouting_Type, 
                    values = c("Combination_Sprouter" = "blue",
                                "Lignotuber_Only"="magenta",
                                "Obligate_Seeder"="brown")) + scale_x_discrete(limits = c("Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only"), labels = c("Combination Resprouter", "Obligate Seeder", "Basal-Only Resprouter")) + labs(y="Fruit Mass (mg)", x = "Post-Fire Regeneration Strategy") + scale_y_continuous(trans = "log10", breaks = c(50, 500, 5000)) +
  coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))

ggplot(data = eucs.data, 
         aes(x = Sprouting_Type, y = seed_mass_mg, fill = Sprouting_Type)) +
  geom_point(aes(y = seed_mass_mg, color = Sprouting_Type), position = position_jitter(width = .15, seed =1), size = 1.5, alpha = 0.6) + scale_color_manual(values = c("Combination_Sprouter" = "blue",
                                 "Lignotuber_Only"="magenta",
                                 "Obligate_Seeder"="brown")) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 3.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + 
  scale_fill_manual(breaks = eucs.data$Sprouting_Type, 
                    values = c("Combination_Sprouter" = "blue",
                                "Lignotuber_Only"="magenta",
                                "Obligate_Seeder"="brown")) + scale_x_discrete(limits = c("Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only"), labels = c("Combination Resprouter", "Obligate Seeder", "Basal-Only Resprouter")) + labs(y="Seed Mass (mg)", x = "Post-Fire Regeneration Strategy") + scale_y_continuous(trans = "log10", breaks = c(0.5, 5, 50)) +
  coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) 
dev.off()
```
#### For testing what axis labels on log scale look most natural

```{r}
ggplot(data = eucs.data, 
         aes(x = Sprouting_Type, y = max_height_m, fill = Sprouting_Type)) +
   geom_point(aes(y = max_height_m, color = Sprouting_Type), position = position_jitter(width = .15, seed =1), size = 1.5, alpha = 0.6) + scale_color_manual(values = c("Combination_Sprouter" = "blue",
                                 "Lignotuber_Only"="magenta",
                                 "Obligate_Seeder"="brown")) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 3.25) + 
  guides(fill = FALSE) + 
  guides(color = FALSE) + 
  scale_fill_manual(breaks = eucs.data$Sprouting_Type, 
                    values = c("Combination_Sprouter" = "blue",
                                "Lignotuber_Only"="magenta",
                                "Obligate_Seeder"="brown")) + 
  scale_x_discrete(limits = c("Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only"), labels = c("Combination Resprouter", "Obligate Seeder", "Basal-Only Resprouter")) + labs(y="Maximum Height (m)", x = "Post-Fire Regeneration Strategy") + scale_y_continuous(trans = "log10", breaks = c(1, 3, 10, 30, 100)) +coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))
```

### For presentation
```{r}
ggplot(data = eucs.scaled.data, 
         aes(x = Sprouting_Type, y = max_height_m, fill = Sprouting_Type)) +
  geom_point(aes(y = max_height_m, color = Sprouting_Type), position = position_jitter(width = .15, seed =1), size = 1.5, alpha = 0.6) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 3.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + scale_x_discrete(limits = c("Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only")) + labs(y="Maximum Height", x = "Post-Fire Regeneration Strategy") + theme(axis.text.y = element_text(size = 20), axis.title.x = element_text(size = 20)) +
  coord_flip() 


ggplot(data = eucs.scaled.data, 
         aes(x = Sprouting_Type, y = relative_bt_by_girth, fill = Sprouting_Type)) +
  geom_point(aes(y = relative_bt_by_girth, color = Sprouting_Type), position = position_jitter(width = .15, seed =1), size = 1.5, alpha = 0.6) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 3.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + scale_x_discrete(limits = c("Combination_Sprouter", "Obligate_Seeder", "Lignotuber_Only")) + theme(axis.text.y = element_text(size = 20), axis.title.x = element_text(size = 20)) + labs(y="Relative Bark Thickness by Girth", x = "Post-Fire Regeneration Strategy") +
  coord_flip() 


theme(axis.text.y = element_text(size = 20), axis.title.x = element_text(size = 20)) +labs(y= "Pagels Lambda")
```



## Including Stem-Only Resprouters


### For Paper

```{r}
#pdf("../Output_figures/MH_boxplots.pdf")
ggplot(data = eucs.data, 
         aes(x = Sprouting_Type3, y = max_height_m, fill = Sprouting_Type3)) +
  geom_point(aes(y = max_height_m, color = Sprouting_Type3), position = position_jitter(width = .15, seed =1), size = 1.5, alpha = 0.6) + scale_color_manual(values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 3.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + 
  scale_fill_manual(breaks = eucs.data$Sprouting_Type, 
                    values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) + scale_x_discrete(limits = c("Combination_Sprouter", "Stem_Only", "Obligate_Seeder", "Lignotuber_Only"), labels = c("Combination\nResprouter", "Stem-Only\nResprouter", "Obligate\nSeeder", "Basal-Only\nResprouter")) + scale_y_continuous(trans = "log10", breaks = c(1, 3, 10, 30, 100)) +coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15, angle = 45)) + labs(x="", y = "Maximum Height (m)")
#dev.off()

#pdf("../Output_figures/RBT_boxplots.pdf")
ggplot(data = eucs.data, 
         aes(x = Sprouting_Type3, y = relative_bt_by_girth, fill = Sprouting_Type3)) +
  geom_point(aes(y = relative_bt_by_girth, color = Sprouting_Type3), position = position_jitter(width = .15, seed =1), size = 1.5, alpha = 0.6) + scale_color_manual(values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 3.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + 
  scale_fill_manual(breaks = eucs.data$Sprouting_Type, 
                    values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) + scale_x_discrete(limits = c("Combination_Sprouter", "Stem_Only", "Obligate_Seeder", "Lignotuber_Only"), labels = c("Combination\nResprouter", "Stem-Only\nResprouter", "Obligate\nSeeder", "Basal-Only\nResprouter")) + scale_y_continuous(trans = "log10", breaks = c(5, 10, 20, 40)) +
  coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15, angle = 45)) + labs(y = "Relative Bark Thickness/100cm Girth", x = "")
#dev.off()

#pdf("../Output_figures/SLA_boxplots.pdf")
ggplot(data = eucs.data, 
         aes(x = Sprouting_Type3, y = sla_mm2_per_mg, fill = Sprouting_Type3)) +
  geom_point(aes(y = sla_mm2_per_mg, color = Sprouting_Type3), position = position_jitter(width = .15, seed =1), size = 1.5, alpha = 0.6) + scale_color_manual(values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 3.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + 
  scale_fill_manual(breaks = eucs.data$Sprouting_Type, 
                    values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) + scale_x_discrete(limits = c("Combination_Sprouter", "Stem_Only", "Obligate_Seeder", "Lignotuber_Only"), labels = c("Combination\nResprouter", "Stem-Only\nResprouter", "Obligate\nSeeder", "Basal-Only\nResprouter")) + scale_y_continuous(trans = "log10", breaks = c(2, 4, 8)) +
  coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15, angle = 45), axis.title.x = element_text(size = 15)) + labs(y = "SLA (mm2/mg)", x = "")
#dev.off()

#pdf("../Output_figures/FWW_boxplots.pdf")
ggplot(data = eucs.data, 
         aes(x = Sprouting_Type3, y = fruit_wall_width_mm, fill = Sprouting_Type3)) +
  geom_point(aes(y = fruit_wall_width_mm, color = Sprouting_Type3), position = position_jitter(width = .15, seed =1), size = 1.5, alpha = 0.6) + scale_color_manual(values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 3.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + 
  scale_fill_manual(breaks = eucs.data$Sprouting_Type, 
                    values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) + scale_x_discrete(limits = c("Combination_Sprouter", "Stem_Only", "Obligate_Seeder", "Lignotuber_Only"), labels = c("Combination\nResprouter", "Stem-Only\nResprouter", "Obligate\nSeeder", "Basal-Only\nResprouter")) + scale_y_continuous(trans = "log10", breaks = c(0.5, 1, 2, 3, 4)) +
  coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15, angle = 45), axis.title.x = element_text(size = 15)) + labs(y = "Fruit Wall Width (mm)", x = "")
#dev.off()




# All on same plot:


MH.boxplot.paper <- ggplot(data = eucs.data, 
         aes(x = Sprouting_Type3, y = max_height_m, fill = Sprouting_Type3)) +
  geom_point(aes(y = max_height_m, color = Sprouting_Type3), position = position_jitter(width = .15, seed =1), size = 2, alpha = 0.6) + scale_color_manual(values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 3.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + 
  scale_fill_manual(breaks = eucs.data$Sprouting_Type, 
                    values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) + scale_x_discrete(limits = c("Combination_Sprouter", "Stem_Only", "Obligate_Seeder", "Lignotuber_Only"), labels = c("Combination\nResprouter", "Stem-Only\nResprouter", "Obligate\nSeeder", "Basal-Only\nResprouter")) + scale_y_continuous(trans = "log10", breaks = c(1, 3, 10, 30, 100)) +coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.text.x = element_text(size = 25), axis.text.y = element_text(size = 25), axis.title.x = element_text(size = 25, vjust = -1.5), plot.margin = margin(l=0.1,b=2,1,1, "cm")) + labs(y="Maximum Height (m)", x = "")
#, hjust = c(0, 0.7, 0) for moving the ctaegory labels a tad to the left


RBT.boxplot.paper <- ggplot(data = eucs.data, 
         aes(x = Sprouting_Type3, y = relative_bt_by_girth, fill = Sprouting_Type3)) +
  geom_point(aes(y = relative_bt_by_girth, color = Sprouting_Type3), position = position_jitter(width = .15, seed =1), size =2, alpha = 0.6) + scale_color_manual(values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 3.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + 
  scale_fill_manual(breaks = eucs.data$Sprouting_Type, 
                    values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) + scale_x_discrete(limits = c("Combination_Sprouter", "Stem_Only", "Obligate_Seeder", "Lignotuber_Only"), labels = c("", "", "", "")) + scale_y_continuous(trans = "log10", breaks = c(5, 10, 20, 40)) +
  coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.text.x = element_text(size = 25), axis.text.y = element_text(size = 25), axis.title.x = element_text(size=25, vjust = -1.5), plot.margin = margin(1,b=2,1,1, "cm")) + labs(y = "Relative Bark Thickness (mm)", x = "")

SLA.boxplot.paper <- ggplot(data = eucs.data, 
         aes(x = Sprouting_Type3, y = sla_mm2_per_mg, fill = Sprouting_Type3)) +
  geom_point(aes(y = sla_mm2_per_mg, color = Sprouting_Type3), position = position_jitter(width = .15, seed =1), size = 2, alpha = 0.6) + scale_color_manual(values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 3.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + 
  scale_fill_manual(breaks = eucs.data$Sprouting_Type, 
                    values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) + scale_x_discrete(limits = c("Combination_Sprouter", "Stem_Only", "Obligate_Seeder", "Lignotuber_Only"), labels = c("", "", "", "")) + scale_y_continuous(trans = "log10", breaks = c(2, 4, 8)) +
  coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.text.x = element_text(size = 25), axis.text.y = element_text(size = 25), axis.title.x = element_text(size = 25, vjust = -1.5), plot.margin = margin(1,b=2,1,1, "cm")) + labs(y = "SLA (mm2/mg)", x = "") 

LA.boxplot.paper <- ggplot(data = eucs.data, 
         aes(x = Sprouting_Type3, y = leaf_area_cm2, fill = Sprouting_Type3)) +
  geom_point(aes(y = leaf_area_cm2, color = Sprouting_Type3), position = position_jitter(width = .15, seed =1), size = 2, alpha = 0.6) + scale_color_manual(values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) + stat_boxplot(geom ='errorbar', width = 0.25) +
  expand_limits(x = 3.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + 
  scale_fill_manual(breaks = eucs.data$Sprouting_Type, 
                    values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) + scale_x_discrete(limits = c("Combination_Sprouter", "Stem_Only", "Obligate_Seeder", "Lignotuber_Only"), labels = c("", "", "", "")) + scale_y_continuous(trans = "log10", breaks = c(3,10,30,100)) +
  coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.text.x = element_text(size = 25), axis.text.y = element_text(size = 25), axis.title.x = element_text(size = 25, vjust = -1.5), plot.margin = margin(1,b=2,1,1, "cm")) + labs(y = "Leaf Area (cm2)", x = "")

pdf("../Output_figures/multi_boxplots.pdf", width = 30, height = 9)
grid.arrange(MH.boxplot.paper, RBT.boxplot.paper, SLA.boxplot.paper, LA.boxplot.paper, nrow = 1)
dev.off()
```


### For Supp. Mat.

We've decided that we would like to have all boxplots in the supps even though some of them have allready been seen in fig 6 in the main text.

```{r}
suppmat.mh <- ggplot(data = eucs.data, 
         aes(x = Sprouting_Type3, y = max_height_m, fill = Sprouting_Type3)) +
  geom_point(aes(y = max_height_m, color = Sprouting_Type3), position = position_jitter(width = .15, seed =1), size = 2, alpha = 0.6) + scale_color_manual(values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) +
  expand_limits(x = 5.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + 
  scale_fill_manual(values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) + scale_x_discrete(limits = c("Combination_Sprouter", "Stem_Only", "Obligate_Seeder", "Lignotuber_Only"), labels = c("Combination\nResprouter", "Stem-Only\nResprouter", "Obligate\nSeeder", "Basal-Only\nResprouter")) + scale_y_continuous(trans = "log10", breaks = c(1, 3, 10, 30, 100)) + coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.text.x = element_text(size = 25), axis.text.y = element_text(size = 25), axis.title.x = element_text(size = 25, vjust = -1.5), plot.margin = margin(l=0.1,b=2,1,1, "cm")) + labs(y="Maximum Height (m)", x = "")

suppmat.rh <- ggplot(data = eucs.data, 
         aes(x = Sprouting_Type3, y = relative_height_by_girth, fill = Sprouting_Type3)) +
  geom_point(aes(y = relative_height_by_girth, color = Sprouting_Type3), position = position_jitter(width = .15, seed =1), size = 2, alpha = 0.6) + scale_color_manual(values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) +
  expand_limits(x = 5.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + 
  scale_fill_manual(values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) + scale_x_discrete(limits = c("Combination_Sprouter", "Stem_Only", "Obligate_Seeder", "Lignotuber_Only"), labels = c("Combination\nResprouter", "Stem-Only\nResprouter", "Obligate\nSeeder", "Basal-Only\nResprouter")) + scale_y_continuous(trans = "log10", breaks = c(6, 10, 20, 30)) + coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.text.x = element_text(size = 25), axis.text.y = element_text(size = 25), axis.title.x = element_text(size = 25, vjust = -1.5), plot.margin = margin(l=0.1,b=2,1,1, "cm")) + labs(y="Relative Height (m)", x = "")

suppmat.sd <- ggplot(data = eucs.data, 
         aes(x = Sprouting_Type3, y = stem_density_g_per_ml, fill = Sprouting_Type3)) +
  geom_point(aes(y = stem_density_g_per_ml, color = Sprouting_Type3), position = position_jitter(width = .15, seed =1), size = 2, alpha = 0.6) + scale_color_manual(values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) +
  expand_limits(x = 5.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + 
  scale_fill_manual(values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) + scale_x_discrete(limits = c("Combination_Sprouter", "Stem_Only", "Obligate_Seeder", "Lignotuber_Only"), labels = c("Combination\nResprouter", "Stem-Only\nResprouter", "Obligate\nSeeder", "Basal-Only\nResprouter")) + scale_y_continuous(trans = "log10", breaks = c(0.51,0.55, 0.60, 0.70, 0.80)) + coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.text.x = element_text(size = 25), axis.text.y = element_text(size = 25), axis.title.x = element_text(size = 25, vjust = -1.5), plot.margin = margin(l=0.1,b=2,1,1, "cm")) + labs(y="Stem Sapwood Density (g/mL)", x = "")

suppmat.rbt <- ggplot(data = eucs.data, 
         aes(x = Sprouting_Type3, y = relative_bt_by_girth, fill = Sprouting_Type3)) +
  geom_point(aes(y = relative_bt_by_girth, color = Sprouting_Type3), position = position_jitter(width = .15, seed =1), size = 2, alpha = 0.6) + scale_color_manual(values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) +
  expand_limits(x = 5.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + 
  scale_fill_manual(values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) + scale_x_discrete(limits = c("Combination_Sprouter", "Stem_Only", "Obligate_Seeder", "Lignotuber_Only"), labels = c("Combination\nResprouter", "Stem-Only\nResprouter", "Obligate\nSeeder", "Basal-Only\nResprouter")) + scale_y_continuous(trans = "log10", breaks = c(2, 5, 10, 20, 40)) + coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.text.x = element_text(size = 25), axis.text.y = element_text(size = 25), axis.title.x = element_text(size = 25, vjust = -1.5), plot.margin = margin(l=0.1,b=2,1,1, "cm")) + labs(y="Relative Bark Thickness (mm)", x = "")

suppmat.sla <- ggplot(data = eucs.data, 
         aes(x = Sprouting_Type3, y = sla_mm2_per_mg, fill = Sprouting_Type3)) +
  geom_point(aes(y = sla_mm2_per_mg, color = Sprouting_Type3), position = position_jitter(width = .15, seed =1), size = 2, alpha = 0.6) + scale_color_manual(values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) +
  expand_limits(x = 5.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + 
  scale_fill_manual(values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) + scale_x_discrete(limits = c("Combination_Sprouter", "Stem_Only", "Obligate_Seeder", "Lignotuber_Only"), labels = c("Combination\nResprouter", "Stem-Only\nResprouter", "Obligate\nSeeder", "Basal-Only\nResprouter")) + scale_y_continuous(trans = "log10", breaks = c(2,4,8)) + coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.text.x = element_text(size = 25), axis.text.y = element_text(size = 25), axis.title.x = element_text(size = 25, vjust = -1.5), plot.margin = margin(l=0.1,b=2,1,1, "cm")) + labs(y="SLA (mm2/mg)", x = "")

suppmat.la <- ggplot(data = eucs.data, 
         aes(x = Sprouting_Type3, y = leaf_area_cm2, fill = Sprouting_Type3)) +
  geom_point(aes(y = leaf_area_cm2, color = Sprouting_Type3), position = position_jitter(width = .15, seed =1), size = 2, alpha = 0.6) + scale_color_manual(values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) +
  expand_limits(x = 5.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + 
  scale_fill_manual(values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) + scale_x_discrete(limits = c("Combination_Sprouter", "Stem_Only", "Obligate_Seeder", "Lignotuber_Only"), labels = c("Combination\nResprouter", "Stem-Only\nResprouter", "Obligate\nSeeder", "Basal-Only\nResprouter")) + scale_y_continuous(trans = "log10", breaks = c(3,10,30,100)) + coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.text.x = element_text(size = 25), axis.text.y = element_text(size = 25), axis.title.x = element_text(size = 25, vjust = -1.5), plot.margin = margin(l=0.1,b=2,1,1, "cm")) + labs(y="Leaf Area (cm2)", x = "")

suppmat.lm <- ggplot(data = eucs.data, 
         aes(x = Sprouting_Type3, y = leaf_mass_g, fill = Sprouting_Type3)) +
  geom_point(aes(y = leaf_mass_g, color = Sprouting_Type3), position = position_jitter(width = .15, seed =1), size = 2, alpha = 0.6) + scale_color_manual(values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) +
  expand_limits(x = 5.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + 
  scale_fill_manual(values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) + scale_x_discrete(limits = c("Combination_Sprouter", "Stem_Only", "Obligate_Seeder", "Lignotuber_Only"), labels = c("Combination\nResprouter", "Stem-Only\nResprouter", "Obligate\nSeeder", "Basal-Only\nResprouter")) + scale_y_continuous(trans = "log10", breaks = c(0.1, 0.3, 1, 3)) + coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.text.x = element_text(size = 25), axis.text.y = element_text(size = 25), axis.title.x = element_text(size = 25, vjust = -1.5), plot.margin = margin(l=0.1,b=2,1,1, "cm")) + labs(y="Leaf Mass (g)", x = "")

suppmat.fww <- ggplot(data = eucs.data, 
         aes(x = Sprouting_Type3, y = fruit_wall_width_mm, fill = Sprouting_Type3)) +
  geom_point(aes(y = fruit_wall_width_mm, color = Sprouting_Type3), position = position_jitter(width = .15, seed =1), size = 2, alpha = 0.6) + scale_color_manual(values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) +
  expand_limits(x = 5.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + 
  scale_fill_manual(values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) + scale_x_discrete(limits = c("Combination_Sprouter", "Stem_Only", "Obligate_Seeder", "Lignotuber_Only"), labels = c("Combination\nResprouter", "Stem-Only\nResprouter", "Obligate\nSeeder", "Basal-Only\nResprouter")) + scale_y_continuous(trans = "log10", breaks = c(0.5, 1, 2, 4)) + coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.text.x = element_text(size = 25), axis.text.y = element_text(size = 25), axis.title.x = element_text(size = 25, vjust = -1.5), plot.margin = margin(l=0.1,b=2,1,1, "cm")) + labs(y="Fruit Wall Width (mm)", x = "")

suppmat.fm <- ggplot(data = eucs.data, 
         aes(x = Sprouting_Type3, y = fruit_mass_mg, fill = Sprouting_Type3)) +
  geom_point(aes(y = fruit_mass_mg, color = Sprouting_Type3), position = position_jitter(width = .15, seed =1), size = 2, alpha = 0.6) + scale_color_manual(values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) +
  expand_limits(x = 5.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + 
  scale_fill_manual(values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) + scale_x_discrete(limits = c("Combination_Sprouter", "Stem_Only", "Obligate_Seeder", "Lignotuber_Only"), labels = c("Combination\nResprouter", "Stem-Only\nResprouter", "Obligate\nSeeder", "Basal-Only\nResprouter")) + scale_y_continuous(trans = "log10", breaks = c(50, 500, 5000)) + coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.text.x = element_text(size = 25), axis.text.y = element_text(size = 25), axis.title.x = element_text(size = 25, vjust = -1.5), plot.margin = margin(l=0.1,b=2,1,1, "cm")) + labs(y="Fruit Mass (mg)", x = "")

suppmat.sm <- ggplot(data = eucs.data, 
         aes(x = Sprouting_Type3, y = seed_mass_mg, fill = Sprouting_Type3)) +
  geom_point(aes(y = seed_mass_mg, color = Sprouting_Type3), position = position_jitter(width = .15, seed =1), size = 2, alpha = 0.6) + scale_color_manual(values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) +
  geom_boxplot(width = .6, outlier.shape = NA, alpha = 0.15) +
  expand_limits(x = 5.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) + 
  scale_fill_manual(values = c("Combination_Sprouter" = "#CC4778FF",
                              "Stem_Only" = "#F89540FF",   
                              "Lignotuber_Only"="#7E03A8FF",
                              "Obligate_Seeder"="#F8E125FF")) + scale_x_discrete(limits = c("Combination_Sprouter", "Stem_Only", "Obligate_Seeder", "Lignotuber_Only"), labels = c("Combination\nResprouter", "Stem-Only\nResprouter", "Obligate\nSeeder", "Basal-Only\nResprouter")) + scale_y_continuous(trans = "log10", breaks = c(0.5, 5, 50)) + coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.text.x = element_text(size = 25), axis.text.y = element_text(size = 25), axis.title.x = element_text(size = 25, vjust = -1.5), plot.margin = margin(l=0.1,b=2,1,1, "cm")) + labs(y="Seed Mass (mg)", x = "")


pdf('../Output_figures/suppmat_boxplots.pdf', width = 21, height = 30)
grid.arrange(suppmat.mh, suppmat.rh, suppmat.sd, suppmat.rbt, suppmat.sla, suppmat.la, suppmat.lm, suppmat.fww, suppmat.fm, suppmat.sm, nrow = 4)
dev.off()
```


